---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup
```{r init, message = FALSE}
library(GimmmeMySC)
library(RColorBrewer)
library(rlist)
library(Seurat)
library(cowplot)
library(GimmeMyPlot)
library(scDblFinder)

set.seed(1234)
options(future.globals.maxSize = 2000 * 1024^2)
options(ggrepel.max.overlaps = 1000)

path <- file.path(
  dirname(getwd()),
  "inst",
  "extdata"
)

SCT <- TRUE
max_dim <- 100
n_dim <- 30
assay <- "RNA"
species <- "human"

l_samples <- list.files(path) %>%  
  .[sapply(., function(x) dir.exists(file.path(path, x)))]  %>%
  .[str_detect(., "KO")]

if (species == "human") {
    func_format <- identity
} else {
    func_format <- str_to_sentence
}

pal_sc <- c(
  rev(brewer.pal(9, "Blues")[-c(1, 3, 5, 7, 9)]),
  # "white",
  "grey90",
  brewer.pal(9, "Reds")[-c(1, 3, 5, 7, 9)]
)
```

# Load data
```{r data}
list_seurat <-  list.map(
    l_samples,
    f(x) ~{
      count_matrix <- file.path(path, x, "outs") %>%
        list.files(full.names =TRUE) %>% 
        .[str_detect(., "filtered_feature_bc_matrix.h5")]
      tmp <- Read10X_h5(count_matrix) %>%
      CreateSeuratObject()
    tmp[["percent_mitochondrial"]] <- PercentageFeatureSet(tmp, pattern = "^((MT)|(mt))-")
    tmp
    }
)
```

```{r data}
types  <- str_remove_all(l_samples, "_?\\d+")

col_inds <- list(
  c("#FB9A99", "#E31A1C"),
  c("#B2DF8A", "#33A02C")
)
color_types <- sapply(col_inds, last)

col_inds <- list.mapv(unique(types), f(i, j) ~colorRampPalette(col_inds[[j]])(length(types[types == i]))) %>% as.character()

seurat <- merge(
  x = list_seurat[[1]],
  y = list_seurat[-1],
  add.cell.ids = paste0(
      types,
      "_",
      # str_remove_all(l_samples, types) %>% str_remove_all("_")
      l_samples
    )
)

seurat$sample <- rownames(seurat[[]])
seurat[[]] <- separate(
  seurat[[]], 
  col = "sample", 
  into = c("Type", "Patient", "Barcode"),
  sep = "_"
)
```

# QC
## Filter cells with low or too high genes
```{r qc_plot}
print_sc_stats(seurat, probs = c(0, 0.1, 0.5, 0.9, 1)) %>%
  kable_sc(digits = 3)

# plot_sc_cor(seurat)

list.map(
  c(paste0("nFeature", "_", assay), "percent_mitochondrial"),
  f(i, j) ~ {
    density_scatter(
      seurat,
      x = i,
      y = paste0("nCount", "_", assay),
      log_x = TRUE,
      log_y = TRUE
  )
  }
    
) %>% plot_grid(plotlist = ., align = "hv")

Idents(seurat) <- "Patient"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10)
)
```

```{r}
seurat_old <- seurat
list_seurat0 <- list_seurat
features <- c("nFeature_RNA", "nCount_RNA", "percent_mitochondrial")
list_seurat <- list.map(
    list_seurat,
    f(i) ~{
      stats_qc <- sapply(
        features,
        function(x) quantile(
            i[[x]], 
            probs = c(0, .025, .05, .1, seq(.25, .75, .25), .9, .95, .975, 1), 
            na.rm = TRUE)
    )
    subset(
      i, 
      subset = 
        nFeature_RNA > stats_qc["10%", "nFeature_RNA"] &
        nFeature_RNA < stats_qc["90%", "nFeature_RNA"] & 
        percent_mitochondrial < stats_qc["90%", "percent_mitochondrial"] & 
        nCount_RNA < stats_qc["90%", "nCount_RNA"]
  )
  }
)

list.mapv(
  seq_along(list_seurat0),
  f(i) ~ print_filtered_cells(list_seurat0[[i]], list_seurat[[i]])
)

seurat <- integrate_multisamples(list_seurat, paste0(types, "_", l_samples))
```

## Doublets
```{r}
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
sce <- scDblFinder(
  GetAssayData(seurat[["RNA"]], slot = "counts"),
  samples = seurat@active.ident,
  aggregateFeatures = FALSE,
  k = NULL,
  processing = "normFeatures",
  dims = 20
)
table(sce$scDblFinder.class)

seurat$doublets <- sce[, row.names(seurat[[]])]$scDblFinder.class

df_doublets <- list.map(
    unique(seurat$Patient),
    f(x) ~ filter(seurat[[]], Patient == x) %>% pull(doublets) %>% table()
) %>% 
    list.rbind() %>%
    as.data.frame() %>%
    rbind(".Total" = colSums(.))

t(df_doublets) %>% 
  plot_bar_2cat(
    count = TRUE,
    stats = FALSE,
    colour = brewer.pal(3, "Set1")[c(1, 3)],
    colour_text = "black"
  )
```

```{r}
# visualize doublets
# plot_dim(
#    seurat, 
#   group.by = "doubletFinder"
# )

# number of singlets and doublets
Idents(seurat) <- "doublets"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10),
    cols = palette_discrete()
)
Idents(seurat) <- "Patient"
seurat <- subset(seurat, subset = doublets == "singlet")
```

## QC after
```{r}
print_sc_stats(seurat, probs = c(0, 0.1, 0.5, 0.9, 1)) %>%
  kable_sc(digits = 3)

list.map(
  c(paste0("nFeature", "_", assay), "percent_mitochondrial"),
  f(i, j) ~ {
    density_scatter(
      seurat,
      x = i,
      y = paste0("nCount", "_", assay),
      log_x = TRUE,
      log_y = TRUE
  )
  }
    
) %>% plot_grid(plotlist = ., align = "hv")

Idents(seurat) <- "Patient"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10)
)
```

```{r qc_sub}
(ncol(seurat) / ncol(seurat_old) * 100 - 100) %>% `*`(-1) %>% round(1) %>% paste("%")

list.map(
    l_samples,
    f(x) ~ list.map(
      c(seurat_old, seurat),
      f(i) ~ filter(i[[]], Patient == x) %>% nrow()
      ) %>% 
    list.rbind()
) %>% 
    list.cbind() %>%
    as.data.frame() %>% 
    set_colnames(l_samples) %>%
    t() %>%
    as.data.frame() %>%
    rbind(`.Total` = colSums(.)) %>%
    mutate(V3 = V1 - V2) %>%
    select(-V1) %>%
    t() %>%
    as.data.frame() %>%
    set_rownames(c("Kept", "Removed"))  %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE
    )
```

## Negative / positive control genes
```{r}
g_neg <- c("DRD1", "ALB", "MYL1", "NPPA") %>% sort()
g_pos <- c("GAPDH", "ACTB", "TMSB4X", "B2M") %>% sort()
g_ctr <- c(g_neg, g_pos)
g_sex <- c("XIST", "DDX3Y")
list.map(
    list(g_sex, g_ctr),
    ~plot_dot(seurat, features = func_format(.), group.by = "Patient")
) %>%
    plot_grid(plotlist = ., ncol = 2, align = "hv")
```

# Gene filtering
```{r filter_mark}
if (!SCT)
    variable_plot(x)
```

# PCA
```{r pca}
s.genes <- cc.genes$s.genes %>% func_format()
g2m.genes <- cc.genes$g2m.genes %>% func_format()

if (!SCT) {
  if (assay == "sketch") {
    seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
  seurat <- ScaleData(seurat)
} else {
  seurat[["RNA"]] <- split(
    seurat[["RNA"]], 
    f = seurat$Patient,
    layers = c("counts","data")
  )
  seurat <- SCTransform(seurat)
  assay <- "SCT"
  DefaultAssay(seurat) <- "SCT"
  if (assay == "sketch") {
     seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
}
seurat <- RunPCA(seurat, reduction_key = "Dim. ", npcs = max_dim, assay = assay)

# Dimension visualisation
plot_dim(
    seurat, 
    # cells = seq(3.2e5),
    reduction = 'pca',
    group.by = "Patient", 
    split.by = "Type", 
    cols = col_inds,
    label = FALSE
)
plot_dim(
  seurat, 
  reduction = 'pca',
  group.by = "Patient", 
  split.by = "Type", 
  cols = color_types,
  label = FALSE
)
# coord <- trans_coord(
#   seurat@reductions$pca@cell.embeddings[, seq(2)], 
#   add = seurat@meta.data[, "seurat_clusters"] %>% paste("Cluster", .)
# )
# plot_dimred(coord, eig, colour = palette_discrete2())
```

## Dimension selection
```{r}
eig <- seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100
p1 <- plot_eig(eig) + 
  geom_hline(yintercept = mean(eig), col = "gray50", lwd = 1, lty = 2) +
  ylab("% Explained\n variance")

p2 <- cumsum(eig) %>% 
  plot_eig()+ 
  ylim(0, NA) +
  ylab("% Cumulated\n explained variance") + 
  geom_hline(yintercept = 80, color = "gray50", lwd =1, lty = 2)

p1 + p2

cumsum(eig) %>% .[. < 80] %>% last()
```

```{r}
DimHeatmap(seurat, dims = seq(15), cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 31:45, cells = 500, balanced = TRUE)
```

```{r}
plot_dim(
    seurat,
    reduction = "pca",
    group.by = "Patient",
    label = FALSE
)
plot_dim(
    seurat,
    reduction = "pca",
    group.by = "Patient",
    split.by = 'Patient',
    label = FALSE
) + NoLegend()
```

```{r}
comp_genes <- list.map(
    seq(max_dim),
    f(x) ~ Loadings(seurat, reduction = "pca") %>%
        as.data.frame() %>%
        select(paste0("PC_", x)) %>%
        arrange(abs(!!sym(paste0("PC_", x)))) %>%
        tail(20) %>%
        rownames()
) %>% set_names(seq(max_dim))

list.map(
  list(g2m.genes, s.genes),
  f(i) ~lapply(comp_genes, function(x) x[x %in% func_format(i)]) %>% compact()
) %>% set_names(c("G2/M", "S"))

# n_dim <- 20
round(cumsum(eig)[n_dim], 1)
```


## Component markers
```{r pca_plot}
# DimHeatmap(seurat, dims = 1)

# Markers
print(seurat[['pca']], dims = 1:5, nfeatures = 5)
VizDimLoadings(seurat, dims = 1:2, reduction = 'pca')

# i <- 1
# seurat[['pca']]@feature.loadings[, i] %>%
#   plot_bar(
#     abs = TRUE,
#     n_max = 10,
#     label_x = "none",
#     geom = "point",
#     title = paste0("Dim ", i, " (", round(seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100, 1) , "%)"),
#     colour = "red"
#   )
```


## Batch effect
```{r umap}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction.key = "Dim. ",
  )
```

```{r umap}
plot_dim(
  seurat,
  group.by = 'Patient',
  label = FALSE
)
plot_dim(
  seurat,
  group.by = 'Patient',
  split.by = 'Patient',
  label = FALSE
) + NoLegend()
```

# Integration
```{r}
integration_method <- "RPCA"
seurat <- IntegrateLayers(
    object = seurat, 
    method = get(paste0(integration_method, "Integration")),
    new.reduction = integration_method,
    # reduction.key = "Dim. ",
    orig.reduction = "pca",
    assay = assay,
    dims.to.integrate = seq(n_dim),
    normalization.method = ifelse(SCT, "SCT", "LogNormalize"),
    verbose = TRUE
)

seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
```

## UMAP
```{r}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction = integration_method
)
```

```{r int}
assay_old <- assay
assay <- "RNA"
p <- plot_feature(seurat, features) %>%
        list.map(f(x, y, z) ~x + ggtitle(str_clean(z)))
plot_grid(plotlist = p, nrow = 2, ncol = 2, align = "hv")
assay <- assay_old
```


```{r}
list.map(
  c("pca", integration_method),
  f(x) ~{
  plot_dim(
    seurat, 
    reduction = x,
    group = "Patient",
    label = FALSE
) %>% theme_sc_dim() + 
          ggtitle(str_to_upper(x))
  }
) %>% plot_grid(plotlist = ., ncol = 2)
```

## Cell cycle
```{r}
seurat <- CellCycleScoring(
  seurat,
  s.features = s.genes, 
  g2m.features = g2m.genes, 
  set.ident = TRUE
)
seurat$CC.Difference <- seurat$S.Score - seurat$G2M.Score

p1 <- plot_feature(
    seurat,
    features = "CC.Difference"
) %>% pluck(1) + 
    labs(title = NULL, color = "S to G2/M")


p2 <- plot_dim(
    object = seurat,
    # cols = palette_continuous(gray = FALSE)(length(levels(Idents(seurat))))
    group.by = "Phase",
    label = FALSE,
    title = TRUE
) +  labs(x = NULL, y = NULL)

p1 + p2

RidgePlot(seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67") %>% func_format(), ncol = 2)
```

## Automatic annotation
### Database processing
```{r}
if (species == "human") {
    hpca <- celldex::HumanPrimaryCellAtlasData()
    dice <- celldex::DatabaseImmuneCellExpressionData()
} else {
    hpca <- celldex::ImmGenData()
    dice <- celldex::MouseRNAseqData()
}

hpca$label.main <- format_annot(hpca$label.main)
dice$label.main <- format_annot(dice$label.main)
hpca$label.fine <- format_annot(hpca$label.fine)
dice$label.fine <- format_annot(dice$label.fine)

bcell_subtypes <- unique(hpca$label.fine) %>%
  .[str_detect(., "FR[A-F]|CLP")] %>%
  sort() %>%
  rev()
is_bcell <- str_starts(hpca$label.fine, "B cell")
keep_cells <- (!is_bcell) | (hpca$label.fine %in% bcell_subtypes)

hpca2 <- hpca[, keep_cells]

cts <- GetAssayData(seurat, layer = 'counts')
```

### Modelling
```{r}
singleR_main <- SingleR(
  test = cts,
  ref = hpca,
  labels = hpca$label.main
)

singleR_fine <- SingleR(
  test = cts,
  ref = hpca2,
  labels = hpca2$label.fine
)

annotation <- list(singleR_main = singleR_main, singleR_fine = singleR_fine)
saveRDS(annotation, file = file.path(path, paste(project,  "_annotation.rds")))
```

### Visualization
```{r}
label_type <- "labels"
x1 <- "singleR_fine"
x <- get(x1)
# x$pruned.labels
# x$scores
x0 <- as.data.frame(x) %>% select(all_of(label_type))
seurat[[]][, x1] <- x0

if (pull(x0, label_type) %>% str_detect("CL\\:\\d*") %>% any()) {
  x0 <- rbind(as.data.frame(colData(hpca.ont)), as.data.frame(colData(dice.ont))) %>%
      filter(str_detect(label.ont, paste0("^", unique(pull(x0, "labels")), "$", collapse = "|"))) %>% 
      filter(!duplicated(label.ont)) %>%
      mutate(labels = label.ont) %>%
      mutate(label.main = label.main %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
    mutate(label.main = label.fine %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("\\:", ", ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
      left_join(x0) %>%
      column_to_rownames()
  label_type <- "label.fine"
}

to_remove <- table(pull(x0, label_type)) %>%
    .[. < lim] %>%
    names()
x0[pull(x0, label_type) %in% to_remove, label_type] <- NA
pull(x0, label_type) %>% 
    cluster_size() %>% 
    kable_sc()
pull(x0, label_type) %>%
    cluster_size() %>% 
  select(n) %>%
  plot_bar(
    label_x = "none",
    label_y = "percent",
    sample_size = nrow(x0),
    digits = 1,
    n_max = 15
  ) + ggtitle(NULL)
if (x1 == "singleR_main") {
seurat$cell_type <- 
  x0[match(rownames(seurat[[]]), rownames(x0)), label_type]
}

to_keep <- paste(paste(unique(seurat$cell_type) %>% na.omit(),  "\\("), collapse = "|")
x0[[label_type]][!str_detect(x0[[label_type]], to_keep)] <- NA

seurat$b_cell <- seurat$cell_subtype
seurat$b_cell[!str_detect(seurat$b_cell, "B cell")] <- NA
seurat$b_cell <- keep_parenthesis(seurat$b_cell)

x0 <- data.frame(labels = seurat$singleR_main)

seurat$cell_type_formatted <- remove_parenthesis(seurat$cell_subtype)
seurat$cell_subtype_formatted <- keep_parenthesis(seurat$cell_subtype)
```

```{r}
plot_dim(
    seurat,
    group.by = "cell_type"
)

plot_dim(
    seurat,
    split.by = "cell_type",
    ncol = 3,
    label = FALSE
) + NoLegend()
```

```{r}
plot_dim(
    seurat,
    group.by = "cell_subtype",
    cols = palette_continuous(gray = FALSE)(unique(seurat$cell_subtype) %>% length())
)

#x11()
list.map(
    unique(seurat$cell_type_formatted) %>% na.omit(),
    f(x) ~subset(seurat, cell_type_formatted == x)
    ) %>%
list.map(
    f(x, y, z) ~
        plot_dim(
            x,
            group.by = "cell_subtype_formatted",
            cols = rep(palette_discrete(), 5)
        ) +
        labs(title = z, x = NULL) 
) %>% plot_grid(plotlist = ., align = "hv", nrow = 2)

cell_categories <- unique(seurat$cell_type_formatted)
cell_list <- map(cell_categories, ~ {
  seurat$cell_subtype[str_detect(seurat$cell_subtype, paste0("^", .x))] %>% na.omit() %>% keep_parenthesis()
}) %>%
  set_names(cell_categories)

p <- list.map(
  compact(cell_list),
  f(x, y, z) ~ cluster_size(x) %>% 
    select(n) %>%
    plot_bar(
        label_x = "none",
        label_y = "percent",
        sample_size = length(x),
        digits = 1,
        n_max = 15,
        ratio = 1
    ) + ggtitle(z)
)

plot_grid(plotlist = p, align = "hv", ncol = 3, nrow = 2)
```

```{r}
zoom_coord <- function(x) x + xlim(-7, 5) + ylim(-7, 8)
plot_dim(
    seurat, 
    group.by = "b_cell"
) %>%
    zoom_coord()

plot_dim(
    seurat,
    split.by = "b_cell",
    # group.by = "b_cell",
    label = FALSE,
    ncol = 3
) %>%
    zoom_coord() +
    NoLegend()
```

```{r}
plotScoreHeatmap(x, labels.use = unique(seurat$cell_subtype), annotation_legend = FALSE)
# plotScoreHeatmap(x, scores.use = 2, annotation_legend = FALSE)
plotDeltaDistribution(x)
```


## Cluster
```{r clus_int}
resolution2 <- seq(.1, .4, 0.05)
resolution <- c(seq(.5, 1, by = 0.1), 1.2, 1.5)
resolution_tot <-c(resolution2, resolution)
DefaultAssay(seurat) <- assay
seurat <- FindNeighbors(
  seurat,
  reduction = integration_method, 
  dims = seq(n_dim)
)

seurat <- FindClusters(
    seurat,
    verbose = FALSE,
    algorithm = 3,
    resolution = resolution_tot,
    graph.name = tail(names(seurat@graphs), 1)
  )

seurat2 <- seurat
prefix_cl <- paste0(assay, "_snn_res.")
cols_to_remove <- paste0(prefix_cl, resolution[-1])
seurat2@meta.data <- seurat2@meta.data[, !colnames(seurat2@meta.data) %in% cols_to_remove]

select(seurat@meta.data, contains(prefix_cl)) %>%
     list.mapv(f(x) ~ levels(x) %>% length()) %>% 
    set_names(names(.) %>% str_remove(prefix_cl)) %>%
  plot_bar(
    label_x = "none",
    label_y = "value",
    sort = "desc",
    digits = 1,
    n_max = 15
  ) + ggtitle(NULL)
```

```{r}
clusters <- grep(prefix_cl, colnames(seurat[[]]), value = TRUE) %>% sort()
clusters2 <- paste0(prefix_cl, resolution2)
list.mapv(clusters, f(i) ~length(unique(seurat@meta.data[, i])))
cl_size <- map_dfr(
    clusters,
    ~ cluster_size(seurat@meta.data[, .x]) %>% 
        pull(1) %>% 
        tibble(
            Clusters = .x,
            cluster = seq_along(.),
            inertia = .
        )
) %>% pivot_wider(names_from = cluster, values_from = inertia, names_prefix = "") %>%
  as.data.frame() %>%
  mutate(Clusters = str_remove_all(Clusters, paste0(assay, "_snn_res.")))
  
cl_size[is.na(cl_size)] <- ""
kable0(cl_size)
```

```{r}
clustree(seurat, prefix = prefix_cl) +
    # scale_color_discrete(type = brewer.pal(length(clusters), "Reds")) +
    scale_color_discrete(type = palette_continuous(gray = FALSE)(length(clusters))) +
    theme(
        legend.title = element_text(face = "italic", size = 15),
        legend.text = element_text(size = 10)
    )
```

```{r}
p_cl <- list.map(
    clusters,
    f(x, y, z) ~
        plot_dim(
            seurat,
            group.by = x,
            cols = palette_continuous()(seurat@meta.data[, x] %>% unique() %>% length())
        ) + 
      NoLegend() + 
      labs(title = str_remove(x, ".*_snn_res."))
)
plot_grid(plotlist = p_cl[seq(8)], align = "hv", ncol = 4, nrow = 2)
plot_grid(plotlist = p_cl[8:15], align = "hv", ncol = 4, nrow = 2)
```

```{r clus_int_res}
best_resolution <- 0.35
snn_cluster <- paste0(
  ifelse(SCT, "SCT", assay),
  "_snn_res.",
  best_resolution
)

Idents(seurat) <- seurat[[]][, snn_cluster] #%>%
  # as.character() %>% 
  # as.numeric() %>% 
  # `+`(1) %>%
  # str_replace_all("^(\\d)$", "0\\1") %>%
  # paste("G", .)

Idents(seurat) %>% 
    cluster_size() %>%
    kable_sc()

# names(cell_type) <- levels(seurat_integrated)
# seurat_integrated <- RenameIdents(seurat_integrated, cell_type)
```


```{r}
plot_dim(
  seurat, 
  cols = palette_continuous()(length(levels(Idents(seurat))))
  )

plot_dim(
    seurat,
    split.by = snn_cluster,
    cols = palette_continuous()(length(levels(Idents(seurat)))),
    label = FALSE,
    ncol = 4
) + NoLegend()

table(Idents(seurat), seurat$cell_subtype_formatted) %>%
    as.data.frame.matrix() %>%
    t() %>%
    as.data.frame() %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        colour =  palette_continuous(gray = FALSE)(length(unique(seurat[[]][, "cell_subtype_formatted"]))),
        pct  = FALSE,
        threshold = 1,
        digits = 1
    ) + labs(y = "% Cells")

tab <- table(Assigned = seurat[[]][, "cell_type"], Clusters = Idents(seurat))
pheatmap(log10(tab + 10), color = colorRampPalette(c('white', brewer.pal(9, "Reds")))(10))

plot_dim(
    seurat,
    split.by = "cell_type",
    ncol = 3
)
```

```{r}
project <- "KO"
saveRDS(seurat, file = file.path(path, paste0(project, ".rds")))
```

```{r}
bcell <- subset(
  seurat, 
  subset = !is.na(seurat$b_cell)
)

saveRDS(bcell, file = file.path(path, paste0(project, "bcell.rds")))
```

## Sample effect
```{r}
table(seurat$Patient)
meta <- seurat@meta.data
meta$cell_id <- rownames(meta)

cells_to_keep <- meta %>%
    group_by(Patient) %>%
    sample_n(size = min(table(seurat$Patient), n()), replace = FALSE) %>%
    pull(cell_id)

seurat_subset <- subset(seurat, cells = cells_to_keep)

plot_dim(
    object = seurat_subset,
    # cols = palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    split.by = "Patient"
)


df_cl <- list.map(
    unique(seurat_subset$Patient),
    f(x) ~ filter(seurat_subset[[]], Patient == x) %>% pull(all_of(snn_cluster)) %>% fct_drop() %>% table()
) %>%
    list.rbind() %>%
    as.data.frame()
    
df_cl %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        colour = col_inds,
        pct  = FALSE,
        digits = 1
    ) + labs(y = "% Cells") +
    geom_hline(yintercept = 50, color = "gray")

df_cl <- list.map(
    unique(seurat$Patient),
    f(x) ~ filter(seurat[[]], Patient == x) %>% pull(all_of(snn_cluster)) %>% fct_drop() %>% table()
) %>%
    list.rbind() %>%
    as.data.frame()
    
df_cl %>%
    t() %>%
    as.data.frame() %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        # colour = palette_discrete(),
        colour = palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
        legend =  colnames(df_cl),
        pct  = FALSE,
        digits = 1
    ) + labs(y = "% Cells")
```

# Markers

## Predictions from dataset
```{r}
library(readxl)
# fileIn <- file.path(path, "41467_2021_27232_MOESM4_ESM.xlsx")
# name <- "symbol"
fileIn <- file.path(path, "B_cell_subet_Marker_Genes-JCB.xlsx")
name <- "Gene name"

l_bcell_markers <- list.map(
    excel_sheets(fileIn),
    f(i) ~ read_excel(fileIn, sheet = i) %>%
        pull(!!sym(name)) %>%
        na.omit() %>%
        func_format()
) 

p <- list.map(
    l_bcell_markers,
  f(i, j, k) ~ {
      tmp <- list(i) %>%
          AddModuleScore(seurat, features = .)
      stats <- quantile(tmp$Cluster1, probs = c(0.01, 0.99))
      tmp <- subset(tmp, subset = Cluster1 >= stats[1] & Cluster1 <= stats[2])
      plot_violin_sc(
        object = tmp,
        features = "Cluster1",
        # idents = c(0:3),
        cols = palette_continuous(gray = FALSE)(length(unique(Idents(tmp))))
      ) %>% 
          pluck(1) +
        labs(title = str_remove_all(k, " ?vs_all"))
      }
)

 plot_grid(
    plotlist = p,
    ncol  = 3,
    align = "hv"
)
```
## Condition markers
```{r feature.heatmaps, eval = FALSE}
FeaturePlot(
  seurat_integrated, 
  features = head(dgea, n = 3) %>% rownames(), 
  split.by = "Patient", 
  min.cutoff = "q10",
  max.cutoff = "q90",
  reduction = 'umap'
) +
    scale_color_gradientn(
        colors = pal_sc
    )

VlnPlot(
    seurat_integrated, 
    features = head(top_genes, 4) %>% rownames(),
    pt.size = 0, 
    group.by = "cell_type",
    split.by = "Patient", 
    split.plot = TRUE,
    ncol = 2
)

library(ComplexHeatmap)
rowAnnotation(
  "HV" = anno_barplot(top_genes$pct.1),
  "AOSD"= anno_barplot(top_genes$pct.2),
  width = unit(10, "cm")
) %>%
  Heatmap(
    top_genes$avg_log2FC,
    name = "Log2FC",
    cluster_rows = TRUE,
    row_labels = rownames(top_genes),
    right_annotation = .,
    width = unit(1, "cm")
    )

theme_set(theme_cowplot())
pop_seurat <- subset(seurat_integrated, idents = "0")
Idents(pop_seurat) <- "Patient"
avg_pop <- as.data.frame(log1p(AverageExpression(pop_seurat, verbose = FALSE)$RNA))
avg_pop$gene <- rownames(avg_pop)
p <- ggplot(avg_pop, aes(S01004A, S01013A)) + 
  geom_point() + 
  ggtitle("0")
LabelPoints(p, points = top_genes %>% head(15) %>% rownames(), repel = TRUE, color = "red")
```

## Session information
```{r end, echo = FALSE}
sessionInfo()
```
