---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup
```{r init, message = FALSE}
set.seed(1234)
library(GimmmeMySC)
options(future.globals.maxSize = 2000 * 1024^2)
options(ggrepel.max.overlaps = 1000)

cwd <- file.path(
  "/shared",
  "projects",
  "immunaid",
  "data"
)

project <- "deep-seq"
path <- file.path(
  cwd,
  "dobino_1124",
  project
)

SCT <- TRUE
max_dim <- 100
n_dim <- 30
assay <- "RNA"

source("utils_theme.R")
source("utils_sc.R")
l_samples <- list.files(path) %>%  
  .[sapply(., function(x) dir.exists(file.path(path, x)))]  %>%
  .[str_detect(., "KO")]
col_inds0 <- GimmeMyPlot:::palette_discrete()[seq_along(l_samples)]

col_pal <- c(
  rev(brewer.pal(9, "Blues")[-c(1, 3, 5, 7, 9)]),
  "white",
  # "grey90",
  brewer.pal(9, "Reds")[-c(1, 3, 5, 7, 9)]
)
```

```{r, eval = FALSE}
seurat <- readRDS(file.path(path, "seurat2.rds"))
```

# Load data
```{r data}
# library(hdf5r)

list_seurat <-  list.map(
    l_samples,
    f(x) ~{
      count_matrix <- file.path(path, x) %>%
        list.files(full.names =TRUE) %>% 
        .[str_detect(., "filtered_feature_bc_matrix.h5")]
      tmp <- Read10X_h5(count_matrix) %>%
      CreateSeuratObject()
    tmp[["percent_mitochondrial"]] <- PercentageFeatureSet(tmp, pattern = "^((MT)|(mt))-")
    tmp
    }
)
```

```{r data}
types  <- str_remove_all(l_samples, "_?\\d+")

col_inds <- list(
  c("#FB9A99", "#E31A1C"),
  c("#B2DF8A", "#33A02C")
)
color_types <- sapply(col_inds, last)

col_inds <- list.mapv(unique(types), f(i, j) ~colorRampPalette(col_inds[[j]])(length(types[types == i]))) %>% as.character()

seurat <- merge(
  x = list_seurat[[1]],
  y = list_seurat[-1],
  add.cell.ids = paste0(
      types,
      "_",
      # str_remove_all(l_samples, types) %>% str_remove_all("_")
      l_samples
    )
)

seurat$sample <- rownames(seurat[[]])
seurat[[]] <- separate(
  seurat[[]], 
  col = "sample", 
  into = c("Type", "Patient", "Barcode"),
  sep = "_"
)
```

# QC
## Filter cells with low or too high genes
```{r qc_plot}
print_sc_stats(seurat, probs = c(0, 0.1, 0.5, 0.9, 1)) %>%
  kable0(digits = 3)
# plot_sc_cor(seurat)

list.map(
  c(paste0("nFeature", "_", assay), "percent_mitochondrial"),
  f(i, j) ~ {
    density_scatter(
      seurat,
      x = i,
      y = paste0("nCount", "_", assay),
      log_x = TRUE,
      log_y = TRUE
  )
  }
    
)%>% plot_grid(plotlist = ., align = "hv")

Idents(seurat) <- "Patient"
plot_sc_violin(
    seurat,
    # features = c(features_atac[-6], "passed_filters"),
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10)
)

# GimmeMyPlot:::plot_cor(
#     seurat@meta.data[, "nCount_RNA"],
#     seurat@meta.data[, "mitochondrial"]
# )
```

```{r}
seurat_old <- seurat
list_seurat0 <- list_seurat
features <- c("nFeature_RNA", "nCount_RNA", "percent_mitochondrial")
list_seurat <- list.map(
    list_seurat,
    f(i) ~{
      stats_qc <- sapply(
        features,
        function(x) quantile(
            i[[x]], 
            probs = c(0, .025, .05, .1, seq(.25, .75, .25), .9, .95, .975, 1), 
            na.rm = TRUE)
    )
    subset(
      i, 
      subset = 
        nFeature_RNA > stats_qc["10%", "nFeature_RNA"] &
        nFeature_RNA < stats_qc["90%", "nFeature_RNA"] & 
        percent_mitochondrial < stats_qc["90%", "percent_mitochondrial"] & 
        nCount_RNA < stats_qc["90%", "nCount_RNA"]
  )
  }
)

list.mapv(
  seq_along(list_seurat0),
  f(i) ~ print_filtered_cells(list_seurat0[[i]], list_seurat[[i]])
)

seurat <- integrate_multisamples(list_seurat, paste0(types, "_", l_samples))
```

## Doublets
```{r}
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
sce <- scDblFinder(
  GetAssayData(seurat[["RNA"]], slot = "counts"),
  samples = seurat@active.ident,
  aggregateFeatures = FALSE,
  k = NULL,
  processing = "normFeatures",
  dims = 20
)
table(sce$scDblFinder.class)

seurat$doublets <- sce[, row.names(seurat[[]])]$scDblFinder.class

df_doublets <- list.map(
    unique(seurat$Patient),
    f(x) ~ filter(seurat[[]], Patient == x) %>% pull(doublets) %>% table()
) %>% 
    list.rbind() %>%
    as.data.frame() %>%
    rbind(".Total" = colSums(.))

t(df_doublets) %>% 
  plot_bar_2cat(
    count = TRUE,
    stats = FALSE,
    colour = brewer.pal(3, "Set1")[c(1, 3)],
    colour_text = "black"
  )
```

```{r}
# visualize doublets
# DimPlot(
#   seurat, 
#   reduction = 'umap.unintegrated', 
#   group.by = "doubletFinder",
#   label = TRUE,
#   repel = TRUE
# )

# number of singlets and doublets

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA", "mitochondrial"),
  pt.size = 0,
  group.by = "doublets",
  ncol = 3
)

seurat <- subset(seurat, subset = doublets == "singlet")
```

## QC after
```{r qc_sub}
(ncol(seurat) / ncol(seurat_old) * 100 - 100) %>% `*`(-1) %>% round(1) %>% paste("%")

list.map(
    l_samples,
    f(x) ~ list.map(
      c(seurat_old, seurat),
      f(i) ~ filter(i[[]], Patient == x) %>% nrow()
      ) %>% 
    list.rbind()
) %>% 
    list.cbind() %>%
    as.data.frame() %>% 
    set_colnames(l_samples) %>%
    t() %>%
    as.data.frame() %>%
    rbind(`.Total` = colSums(.)) %>%
    mutate(V3 = V1 - V2) %>%
    select(-V1) %>%
    t() %>%
    as.data.frame() %>%
    set_rownames(c("Kept", "Removed"))  %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE
    )
```

## Biotype
```{r, eval = FALSE}
# genome_species <- "Hsapiens"
genome_species <- "Mmusculus"
ref_path <- "/shared/projects/immunaid/data/ref"
annotation_file <- file.path(ref_path, paste(tolower(genome_species), 113, "annotation.tsv", sep = "_"))
annotation <- read_tsv(annotation_file) %>% as.data.frame()

gene_query <- mutate(annotation, biotype = gene_biotype %>% 
  str_replace_all(".*(pseudogene)", "Pseudogene") %>%
  str_replace_all("(IG)_.*", "\\1") %>%
  str_replace_all("(TR)_.*", "\\1") %>%
  str_replace_all(".*(decay)", "Decayed") %>%
  str_replace_all("(misc_RNA)|(snRNA)|(miRNA)|(snoRNA)|(scaRNA)|(sRNA)|(scRNA)|(vault_RNA)|(Mt_ncRNA)|(Mt_tRNA)", "sncRNA") %>%
  str_replace_all("Mt_rRNA", "rRNA") %>% 
  # str_replace_all("retained_intron", "lncRNA") %>% 
  str_replace_all("(IG)|(TR)", "somatic recombination") %>% 
  # str_replace_all("(somatic recombination)|(TEC)|(rRNA)|(ribozyme)", "Others") %>% 
  str_replace_all("^protein_coding$", "CDS") %>%
  str_replace_all("processed_transcript", "ncRNA")
)


cts <- GetAssayData(seurat, layer = 'counts')
common_genes <- intersect(pull(gene_query, external_gene_name), rownames(cts))
biotype <- list.map(
  c("CDS", "Pseudogene", "rRNA", "lncRNA", "sncRNA"),
    f(i) ~{
      n <- filter(gene_query, external_gene_name %in% common_genes) %>%
        filter(biotype == y) %>% 
        filter(!duplicated(external_gene_name)) %>%
        nrow()
      list.map(
        rownames(cts),
        f(x) ~{
          tmp <- gene_query %>% filter(external_gene_name == x & biotype == i)
          if (nrow(tmp) > 0) {
            cts0 <- cts %>%
              filter(rownames(.) == x) %>% 
              as.numeric() %>%
              `>`(1) %>% 
              as.numeric()
          } else {
            rep(0, ncol(cts))
          }
        }
      ) %>% list.cbind() %>%
      apply(1, sum) %>% `/`(n) %>% `*`(100)
  }
)
```

## Negative / positive control genes
```{r, eval = FALSE}
g_neg <- c("DRD1", "ALB", "MYL1", "NPPA") %>% sort()
g_pos <- c("GAPDH", "ACTB", "TMSB4X", "B2M") %>% sort()
g_ctr <- c(g_neg, g_pos)
g_sex <- c("XIST", "DDX3Y")
DotPlot(
    seurat, 
    features = str_to_sentence(g_sex), 
    scale = FALSE,
    dot.scale = 8
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL)

list.map(
    list(g_neg, g_pos),
    ~DotPlot(
    seurat, 
    features = str_to_sentence(.) %>% sort(), 
    dot.scale = 8
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL)
)%>%
    plot_grid(plotlist = ., ncol = 2, align = "hv")
```

## Scater

# Gene filtering
```{r filter}
seurat[["RNA"]] <- split(seurat[["RNA"]], f = seurat$Patient)
if (!SCT) {
  seurat <- NormalizeData(seurat) %>%
      FindVariableFeatures()
}
```

```{r eval = FALSE}
# cts0 <- GetAssayData(seurat, layer = 'data')
cts <- as.matrix(seurat[["RNA"]]$counts) %>% as.data.frame()
cts0 <- as.matrix(seurat[["RNA"]]$data) %>% as.data.frame()

n_cell <- 1e3
plot_density_genes(select(cts0, seq(n_cell)), density = TRUE, norm = TRUE)
plot_density_genes(select(cts, seq(n_cell)), density = TRUE)
```

```{r filter, eval = FALSE}
vst <- seurat[["RNA"]]@meta.data[, seq(2)]
vst0 <- seurat[["RNA"]]@meta.data[, 3:4]
# vst0 <- sapply(
#   c(mean, var),
#   function(i) apply(cts, 1, i)
#   ) %>% as.data.frame()

vst %>% 
  # select(vf_vst_counts_mean, vf_vst_counts_variance.standardized) %>%
  set_colnames(c("Average expression", "Variance")) %>%
  ggplot(mapping = aes(x = `Average expression`, y = Variance)) + 
    geom_point() +
    scale_x_continuous(trans = log10_trans(),
      breaks = trans_breaks("log10", function(x) 10^x),
      labels = trans_format("log10", math_format(10^.x))) +
    scale_y_continuous(trans = log2_trans(),
      breaks = trans_breaks("log2", function(x) 2^x),
      labels = trans_format("log2", math_format(2^.x))) +
    theme_custom()
```

```{r filter_mark}
if (!SCT)
VariableFeaturePlot(seurat) %>%
    LabelPoints(
        points = head(VariableFeatures(seurat), 10), 
        repel = TRUE
    ) +
    scale_x_continuous(trans = log10_trans(),
                       breaks = trans_breaks("log10", function(x) 10^x),
                       labels = trans_format("log10", math_format(10^.x))) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = label_number_auto()) +
    theme_custom() +
    ylab("Variance") +
    NoLegend()
```

# PCA
```{r pca}
s.genes <- cc.genes$s.genes %>% str_to_sentence()
g2m.genes <- cc.genes$g2m.genes %>% str_to_sentence()

if (!SCT) {
  if (assay == "sketch") {
    seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
  seurat <- ScaleData(seurat)
} else {
  seurat[["RNA"]] <- split(
    seurat[["RNA"]], 
    f = seurat$Patient,
    layers = c("counts","data")
  )
  seurat <- SCTransform(seurat)
  assay <- "SCT"
  DefaultAssay(seurat) <- "SCT"
  if (assay == "sketch") {
     seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
}
seurat <- RunPCA(seurat, reduction_key = "Dim. ", npcs = max_dim)

# Dimension visualisation
DimPlot(
    seurat, 
    # cells = seq(3.2e5),
    reduction = 'pca',
    group.by = "Patient", 
    split.by = "Type", 
    cols = col_inds,
    repel = TRUE,
    pt.size = .5,
    alpha = .25
) %>% theme_sc_dim()
DimPlot(
  seurat, 
  reduction = 'pca',
  group.by = "Type", 
  cols = color_types
) %>% theme_sc_dim()
# coord <- trans_coord(
#   seurat@reductions$pca@cell.embeddings[, seq(2)], 
#   add = seurat@meta.data[, "seurat_clusters"] %>% paste("Cluster", .)
# )
# plot_dimred(coord, eig, colour = GimmeMyPlot:::palette_discrete2())
```

## Cell cycle
```{r}
seurat <- CellCycleScoring(
  seurat,
  s.features = s.genes, 
  g2m.features = g2m.genes, 
  set.ident = TRUE
)
seurat$CC.Difference <- seurat$S.Score - seurat$G2M.Score

FeaturePlot(
    seurat,
    features = "CC.Difference",
    alpha = 0.5
) %>% theme_sc_dim()  +
    scale_color_gradientn(
        colors = col_pal,
        limits = c(-1, 1),
        name = "S to G2/M"
    ) + labs(x = NULL, y = NULL, title = NULL)


DimPlot(
    object = seurat, 
    label = TRUE,
    # cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    cols = GimmeMyPlot:::palette_discrete(),
    repel = TRUE,
    group.by = "Phase",
    na.value = "white",
    reduction = "umap",
    pt.size = 1,
    alpha = 0.25
) %>% 
    theme_sc_dim() +  labs(x = NULL, y = NULL, title = NULL)

RidgePlot(seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
```

```{r, eval = FALSE}
seurat <- RunPCA(seurat, features = c(s.genes, g2m.genes))
DimPlot(seurat)

cc_diff <- TRUE
if (!cc_diff)
  seurat <- ScaleData(seurat, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(seurat))
else
  seurat <- ScaleData(seurat, vars.to.regress = "CC.Difference", features = rownames(seurat))

seurat <- RunPCA(seurat, features = VariableFeatures(seurat), nfeatures.print = 10)
seurat <- RunPCA(seurat, reduction_key = "Dim. ", npcs = max_dim)
```

## Dimension selection
```{r}
# Dimension selection
plot_eig(seurat[['pca']]@stdev) + 
   geom_hline(yintercept = mean(seurat[['pca']]@stdev), col = "gray50", lwd =1, lty = 2)

eig <- seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100
plot_eig(eig) + 
  geom_hline(yintercept = mean(eig), col = "gray50", lwd = 1, lty = 2) +
  ylab("% Explained\n variance")
# cumsum(eig) %>% .[. < 80]
cumsum(eig) %>% 
  plot_eig()+ 
  ylim(0, NA) +
  ylab("% Cumulated\n explained variance") + 
  geom_hline(yintercept = 80, color = "gray50", lwd =1, lty = 2)
```

```{r}
DimHeatmap(seurat, dims = seq(15), cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 31:45, cells = 500, balanced = TRUE)

DimPlot(
  seurat,
  reduction = "pca",
  group.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  pt.size = .5,
  alpha = .25
) %>% theme_sc_dim() + ggtitle(NULL)
```

```{r}
comp_genes <- list.map(
    seq(max_dim),
    f(x) ~ Loadings(seurat, reduction = "pca") %>%
        as.data.frame() %>%
        select(paste0("PC_", x)) %>%
        arrange(abs(!!sym(paste0("PC_", x)))) %>%
        tail(20) %>%
        rownames()
) %>% set_names(seq(max_dim))

list.map(
  list(g2m.genes, s.genes),
  f(i) ~lapply(comp_genes, function(x) x[x %in% str_to_sentence(i)]) %>% compact()
) %>% set_names(c("G2/M", "S"))

# n_dim <- 20
round(cumsum(eig)[n_dim], 1)
```


## Component markers
```{r pca_plot, eval = FALSE}
# DimHeatmap(seurat, dims = 1)

# Markers
print(seurat[['pca']], dims = 1:5, nfeatures = 5)
VizDimLoadings(seurat, dims = 1:2, reduction = 'pca')

# i <- 1
# seurat[['pca']]@feature.loadings[, i] %>%
#   plot_bar(
#     abs = TRUE,
#     n_max = 10,
#     label_x = "none",
#     geom = "point",
#     title = paste0("Dim ", i, " (", round(seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100, 1) , "%)"),
#     colour = "red"
#   )
```

# Batch effect
## Cluster
```{r clus_unint, eval = FALSE}
resolution <- c(seq(.1, .4, 0.05), seq(.5, 1, by = 0.1), 1.2, 1.5)
seurat <- FindNeighbors(seurat, dims = seq(n_dim)) 
seurat <- seurat %>% 
  FindClusters(verbose = FALSE, algorithm = 3, resolution = resolution)

select(seurat2@meta.data, contains("RNA_snn_res")) %>%
    list.mapv(f(x) ~ levels(x) %>% length())

resolution <- 0.1
Idents(seurat) <- "seurat_clusters"
col_clus <- get_colors0()(length(levels(Idents(seurat))))
```

## UMAP
```{r umap}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction.key = "Dim. ",
  # reduction.name = "umap.unintegrated"
  )

# Batch effect
DimPlot(
  seurat,
  # reduction = 'umap.unintegrated',
  group.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  pt.size = .5,
  alpha = .25
) %>% theme_sc_dim() + ggtitle(NULL)
DimPlot(
  seurat,
  # reduction = 'umap.unintegrated',
  split.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  alpha = .25
) %>% theme_sc_dim() + NoLegend()
```


```{r, eval = FALSE}
DimPlot(
  seurat,
  reduction = 'umap.unintegrated',
  group.by = 'Type',
  cols = color_types,
  repel = TRUE
) %>% theme_sc_dim() +
DimPlot(
  seurat,
  reduction = 'umap.unintegrated',
  group.by = 'seurat_clusters',
  cols = col_clus,
  label = TRUE,
  repel = TRUE
)

DimPlot(
  seurat,
  # reduction = 'umap.unintegrated',
  group.by = "Patient",
  split.by = "Type",
  cols = col_inds,
  repel = TRUE
) %>% theme_sc_dim()
```

# Integration
```{r, eval = FALSE}
if(SCT) {
  seurat[["v5"]] <- CreateAssay5Object(
    counts =  seurat@assays$RNA$counts,
    data =  seurat@assays$SCT$data,
    # scale.data = seurat@assays$SCT$scale.data
  )
  seurat@assays[["v5"]]@key <- "v5_"
  DefaultAssay(seurat) <- "v5"
    seurat[["v5"]] <- split(
    seurat[["v5"]], 
    f = seurat$Patient,
    layers = c("counts","data")
  )
  assay <- "V5"
}
```

```{r}
integration_method <- "RPCA"
seurat <- IntegrateLayers(
    object = seurat, 
    method = get(paste0(integration_method, "Integration")),
    new.reduction = integration_method,
    # reduction.key = "Dim. ",
    orig.reduction = "pca",
    assay = assay,
    dims.to.integrate = n_dim,
    # scale.layer = "data",
    normalization.method = ifelse(SCT, "SCT", "LogNormalize"),
    verbose = TRUE
)
# integration_method <- "RPCA"
# seurat <- IntegrateLayers(
#   object = seurat, 
#   method = get(paste0(integration_method, "Integration")),
#   new.reduction = integration_method,
#   # reduction.key = "Dim. ",
#   # features = Features(seurat@assays[["v5"]]),
#   orig.reduction = "pca",
#   # assay = "v5",
#   dims.to.integrate = n_dim,
#   # scale.layer = "data",
#   normalization.method = ifelse(SCT, "SCT", "LogNormalize"),
#   verbose = TRUE
# )

seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
```

## UMAP
```{r}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction = integration_method
)
```

```{r int}
assay_old <- assay
assay <- "RNA"
p <- map(
    features,
    ~FeaturePlot(
        object = seurat,
        features = .,
        alpha = 0.5,
        # cols = c("yellow", "blue"), 
        # min.cutoff = "q10", 
        # max.cutoff = "q90"
    ) %>% 
        theme_sc_dim() +
        ggtitle(str_clean(.)) +
    scale_color_gradientn(
        colors = col_pal
    )

)

plot_grid(plotlist = p, nrow = 2, ncol = 2, align = "hv")
assay <- assay_old
```


```{r, eval = FALSE}
list.map(
  c("pca", integration_method),
  f(x) ~{
  DimPlot(
    seurat_integrated, 
    reduction = x, 
    group = "Patient",
    cols = col_inds,
    label = TRUE,
    repel = TRUE
) %>% theme_sc_dim() + ggtitle(x)
  }
) %>% plot_grid(plotlist = ., ncol = 2)

DimPlot(
  seurat_integrated, 
  reduction = integration_method, 
  group = "Type",
  cols = color_types,
  repel = TRUE
) %>% theme_sc_dim()
```


```{r}
DimPlot(
  seurat, 
  reduction = integration_method, 
  group = "Patient",
  split.by = "Type",
  cols = col_inds,
  repel = TRUE
) %>% theme_sc_dim()
# seurat[[assay]] <- JoinLayers(seurat[[assay]])
```

## Cluster
```{r clus_int}
resolution2 <- seq(.1, .4, 0.05)
resolution <- c(seq(.5, 1, by = 0.1), 1.2, 1.5)
resolution_tot <-c(resolution2, resolution)
DefaultAssay(seurat) <- assay
seurat <- FindNeighbors(
  seurat,
  reduction = integration_method, 
  dims = seq(n_dim)
)

seurat <- FindClusters(
    seurat,
    verbose = FALSE,
    algorithm = 3,
    resolution = resolution_tot,
    graph.name = tail(names(seurat@graphs), 1)
  )

prefix_cl <- paste0(assay, "_snn_res.")
select(seurat@meta.data, contains(prefix_cl)) %>%
     list.mapv(f(x) ~ levels(x) %>% length())
```

```{r clus_int_res, eval = FALSE}
resolution <- 0.1
snn_cluster <- paste0(
  ifelse(SCT, "SCT", assay),
  "_snn_res.",
  resolution
)
Idents(seurat_integrated) <- snn_cluster
col_clus <- get_colors0()(length(levels(Idents(seurat_integrated))))

DimPlot(
  seurat_integrated, 
  reduction = "pca",
  group.by = snn_cluster,
  cols = col_clus,
  # cols = brewer.pal(12, "Paired"),
  label = TRUE,
  repel = TRUE
  )
```

```{r clus_int_plot, eval = FALSE}
# Look at cluster IDs of the first 5 cells
head(Idents(seurat_integrated), 5)

cell_type <- Idents(seurat_integrated) %>%
  as.character() %>% 
  as.numeric() %>% 
  `+`(1) %>%
  str_replace_all("^(\\d)$", "0\\1") %>%
  paste("Cluster", .)

cell_type %>%
  fct_count() %>%
  column_to_rownames("f") %>%
    mutate(freq = round(n/sum(n) * 100, 1)) # %>%
  # plot_bar(label_x = "value")

# names(cell_type) <- levels(seurat_integrated)
# seurat_integrated <- RenameIdents(seurat_integrated, cell_type)
```


```{r}
stop()
```

```{r}
# Visualization
# DimPlot(seurat_integrated, reduction = 'umap', group.by=c('Patient', 'Type'))
DimPlot(
  seurat_integrated,
  reduction = 'umap',
  group.by = 'Patient',
  cols = col_inds,
  repel = TRUE
) %>% theme_sc_dim() +
DimPlot(
  seurat_integrated,
  reduction = 'umap',
  group.by = 'Type',
  cols = color_types,
  repel = TRUE
) %>% theme_sc_dim() +
DimPlot(
  seurat_integrated,
  reduction = 'umap',
  group.by = snn_cluster,
  cols = col_clus,
  label = TRUE,
  repel = TRUE
)

DimPlot(
  seurat_integrated, 
  reduction = "umap", 
  group = "Patient",
  split.by = "Type",
  cols = col_inds,
  repel = TRUE
) %>% theme_sc_dim()

list.map(
  c("umap.unintegrated", "umap"),
  f(x) ~{
  DimPlot(
    seurat_integrated, 
    reduction = x, 
    group = "Patient",
    cols = col_inds,
    repel = TRUE
) %>% theme_sc_dim() + ggtitle(x)
  }
) %>% plot_grid(plotlist = ., ncol = 2)

list.map(
  c("umap.unintegrated", "umap"),
  f(x) ~{
  DimPlot(
    seurat_integrated, 
    reduction = x, 
    group = "Type",
    cols = color_types,
    repel = TRUE
) %>% theme_sc_dim() + ggtitle(x)
  }
) %>% plot_grid(plotlist = ., ncol = 2)
```

```{r}
SaveSeuratRds(
  seurat_integrated, 
  file = file.path(cwd, "Seurat", "seurat_stillhealthy.rds"), 
  compress = TRUE
)
```

# Markers
## Cell markers
```{r marker_cell}
seurat_integrated$cell_type <- seurat_integrated@meta.data[, snn_cluster] %>%
  as.character() %>% 
  as.numeric() %>%
  `+`(1) %>%
  paste0("G", .)

# makers for epithelial endothelial mesenchymal and immune cells
FeaturePlot(
  seurat_integrated,
  features = c("EPCAM", "CLDN5", "COL1A2", "PTPRC"),
  min.cutoff = "q10", 
  max.cutoff = "q90",
  label = TRUE,
  cols = c("grey", "red")
)

# markers for macrophage, neutrophil, B cell, T cell, NK cell 
FeaturePlot(
  seurat_integrated,
  features = c("AIF1", "S100A8", "BANK1", "CD3G", "NKG7"),
  min.cutoff = "q10", 
  max.cutoff = "q90",
  label = TRUE,
  cols = c("grey", "red")
)

# DefaultAssay(seurat_integrated)
if (SCT)
  seurat_integrated <- PrepSCTFindMarkers(seurat_integrated)

cell_markers <- lapply(
  levels(Idents(seurat_integrated)),
  function(x) {
  FindConservedMarkers(
    seurat_integrated, 
    ident.1 = x, 
    grouping.var = "Patient", 
    verbose = TRUE
  )
  }
)

(top_genes <- lapply(
  cell_markers,
  function(x) {
        head(x, n = 10) %>%
          # filter(abs(avg_log2FC) > 1) %>%
          rownames()
  }
))

# head(cluster2.markers, n = 10) %>% 
#   `^`(2, .) %>%
#   select("avg_log2FC") %>% 
#   plot_bar(label_x = "value", digits = 1, title = "FC")

top_genes2 <- lapply(
  c(1, 2, 5),
  function(i) {
  lapply(top_genes, function(x) head(x,  i)) %>% unlist() %>% sort() %>% unique()
  }
)
RidgePlot(
  seurat_integrated, 
  features = top_genes2[[1]], 
  ncol = 4,
  cols = col_clus
)
VlnPlot(
  seurat_integrated, 
  features = top_genes2[[1]],
  pt.size = 0,
  cols = col_clus
)
VlnPlot(
  seurat_integrated, 
  features = top_genes2[[1]],
  pt.size = 0, 
  split.by = "Patient", 
  split.plot = TRUE
)
# VlnPlot(seurat_integrated, features = top_genes[[1]], slot = 'counts', log = TRUE)
FeaturePlot(
  seurat_integrated,
  features = top_genes2[[1]],
  min.cutoff = "q10", 
  max.cutoff = "q90",
  label = TRUE,
  cols = c("grey", "red")
)

DoHeatmap(
  subset(seurat_integrated, downsample = 100),
  features = top_genes2[[2]],
  group.colors = col_clus,
  group.by = "cell_type"
  ) + scale_fill_gradientn(
    colors = c("blue", "white", "red"), 
    na.value = "black"
  )
DotPlot(
  seurat_integrated, 
  features = top_genes2[[2]], 
  cols = c('blue', 'red'), 
  dot.scale = 8,
  split.by = "Patient"
  ) +
  RotatedAxis()+
  theme_custom()
```

## Automatic annotation
```{r}
# p1 <- DimPlot(
#   seurat_integrated, 
#   reduction = 'umap', 
#   group.by = snn_cluster,
#   cols = col_clus,
#   label = TRUE,
#   repel = TRUE
# )

hpca <- celldex::HumanPrimaryCellAtlasData()
dice <- celldex::DatabaseImmuneCellExpressionData()
monaco <- celldex::MonacoImmuneData()
hpca <- celldex::ImmGenData()
dice <- celldex:::MouseRNAseqData()

hpca$label.main <- format_annot(hpca$label.main)
dice$label.main <- format_annot(dice$label.main)
monaco$label.main <- format_annot(monaco$label.main)
hpca$label.fine <- format_annot(hpca$label.fine)
# hpca2 <- filter(hpca, !str_detect(label.main, "B cell") | str_detect(label.fine, "FR[A-F]|CLP"))
dice$label.fine <- format_annot(dice$label.fine)
monaco$label.fine <- format_annot(monaco$label.fine)

bcell_subtypes <- unique(hpca$label.fine) %>%
  .[str_detect(., "FR[A-F]|CLP")] %>%
  sort() %>%
  rev()
is_bcell <- str_starts(hpca$label.fine, "B cell")
keep_cells <- (!is_bcell) | (hpca$label.fine %in% bcell_subtypes)

hpca2 <- hpca[, keep_cells]
# View(as.data.frame(colData(hpca)))

cts <- GetAssayData(seurat, layer = 'counts')

# seurat_annotation <- list()
# seurat_annotation <- readRDS(file.path(cwd, "Seurat", "seurat_annotation.rds"))
```

```{r, eval = FALSE}
lim <- 100
label_type <- "labels"
x1 <- "singleR_fine"
x <- get(x1)
# x$pruned.labels
# x$scores
x0 <- as.data.frame(x) %>% select(label_type) # %>% mutate(rowname = rownames(.))

seurat[[]][, x1] <- x0
# saveRDS(seurat_annotation, file = file.path(cwd, "Seurat", "seurat_annotation.rds"))

if (pull(x0, label_type) %>% str_detect("CL\\:\\d*") %>% any()) {
  x0 <- rbind(as.data.frame(colData(hpca.ont)), as.data.frame(colData(dice.ont))) %>%
      filter(str_detect(label.ont, paste0("^", unique(pull(x0, "labels")), "$", collapse = "|"))) %>% 
      filter(!duplicated(label.ont)) %>%
      mutate(labels = label.ont) %>%
      mutate(label.main = label.main %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
    mutate(label.main = label.fine %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("\\:", ", ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
      left_join(x0) %>%
      column_to_rownames()
  label_type <- "label.fine"
}

# fct_count(pull(x0, label_type)) %>% arrange(desc(n)) %>% tail(20) %>% print(20)
to_remove <- table(pull(x0, label_type)) %>% .[. < lim] %>% names()
# seurat[[]][, x0] <- x0
x0[pull(x0, label_type) %in% to_remove, label_type] <- NA
pull(x0, label_type) %>% cluster_size() %>% kable0()
pull(x0, label_type) %>% cluster_size() %>% 
  select(n) %>%
  GimmeMyPlot::plot_bar(
    label_x = "none",
    label_y = "percent",
    sample_size = nrow(x0),
    digits = 1,
    n_max = 15
  ) + ggtitle(NULL)
seurat$cell_subtype <- 
  x0[match(rownames(seurat[[]]), rownames(x0)), label_type]
# Idents(seurat) <- x0[match(rownames(seurat[[]]), rownames(x0)), label_type]

to_keep <- paste(paste(unique(seurat$cell_type) %>% na.omit(),  "\\("), collapse = "|")
x0[[label_type]][!str_detect(x0[[label_type]], to_keep)] <- NA
# x0[[label_type]][str_detect(x0[[label_type]], "B cell") & !(x0[[label_type]] %in% bcell_subtypes)] <- NA

seurat$b_cell <- seurat$cell_subtype
seurat$b_cell[!str_detect(seurat$b_cell, "B cell")] <- NA
seurat$b_cell <- keep_parenthesis(seurat$b_cell)
to_remove <- table(seurat$b_cell) %>% .[. < 100] %>% names()
seurat$b_cell[seurat$b_cell %in% to_remove] <- NA

label_type <- "labels"
x0 <- data.frame(labels = seurat$singleR_main)

seurat$id <- remove_parenthesis(seurat$cell_subtype)
seurat$id2 <- keep_parenthesis(seurat$cell_subtype)

Idents(seurat) <- pull(x0, label_type) %>% as.factor()
# Idents(seurat) <- pull(x0, label_type)  %>% str_remove_all( "B cell \\(|\\)") %>% str_replace("(\\w+)\\.(\\w+)", "\\1 (\\2)") %>% str_replace( "FR", "Fr") %>% as.factor()
type <- "B cell"
x2 <- mutate(x0, labels = ifelse(str_detect(labels, type), labels, NA))

type <- NULL
DimPlot(
    object = seurat, 
    label = TRUE,
    # cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    cols = GimmeMyPlot:::palette_discrete(),
    repel = TRUE,
    na.value = "white",
    reduction = "umap",
    pt.size = 1,
    alpha = 0.25
) %>% 
    theme_sc_dim() + ggtitle(type)

type <- "SCT_snn_res.0.35"
table(seurat[[]][, type], seurat$id2) %>%
    as.data.frame.matrix() %>%
    t() %>%
    as.data.frame() %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        colour =  GimmeMyPlot:::palette_continuous(gray = FALSE)(length(unique(seurat[[]][, "id2"]))),
        pct  = FALSE,
        threshold = 1,
        digits = 1
    ) + labs(y = "% Cells")

cell_categories <- seurat$cell_subtype %>%
  str_split(" \\(", simplify = TRUE) %>%
  .[,1] %>%
  unique()


cell_list <- map(cell_categories, ~ {
  seurat$cell_subtype[str_detect(seurat$cell_subtype, paste0("^", .x))] %>% na.omit() %>% keep_parenthesis()
}) %>%
  set_names(cell_categories)

p <- list.map(
  compact(cell_list),
  f(x, y, z) ~ cluster_size(x) %>% 
    select(n) %>%
    GimmeMyPlot::plot_bar(
        label_x = "none",
        label_y = "percent",
        sample_size = length(x),
        digits = 1,
        n_max = 15,
        ratio = 1
    ) + ggtitle(z)
)

plot_grid(plotlist = p, align = "hv", ncol =3, nrow = 2)

x2 <- seurat$singleR_fine
x2[!x2 %in% bcell_subtypes] <- NA
Idents(seurat) <- x2
    
DimPlot(
    seurat, 
    reduction = 'umap', 
    group.by = "cell_type",
    cols = col_cont(unique(seurat$cell_type) %>% length()),
    label = TRUE,
    repel = TRUE,
    alpha = 0.5,
    na.value = "white"
) %>% theme_sc_dim() + ggtitle(NULL)

DimPlot(
    seurat, 
    reduction = 'umap', 
    group.by = "b_cell",
    cols = col_cont(unique(seurat$b_cell) %>% length()),
    label = TRUE,
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white"
) %>% theme_sc_dim() + ggtitle(NULL) + 
    xlim(-7, 5) + ylim(-7, 8)

DimPlot(
    seurat, 
    reduction = 'umap', 
    split.by = "b_cell",
    group.by = "b_cell",
    cols = col_cont(unique(seurat$b_cell) %>% length()),
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white",
    ncol = 3
) %>% theme_sc_dim() + ggtitle(NULL) + 
    xlim(-7, 5) + ylim(-7, 8)

list.map(
    seurat_type,
    f(x, y, z) ~
        DimPlot(
            x, 
            reduction = 'umap.unintegrated',
            group.by = "cell_annotation",
            na.value = "white",
            cols = col_cont(unique(seurat$cell_annotation) %>% length()),
            label = TRUE,
            repel = TRUE
        )  %>% theme_sc_dim() + NoLegend() + ggtitle(z)
) %>% plot_grid(plotlist = ., align = "hv", ncol = 2, nrow = 2)

# plotScoreHeatmap(x)
plotScoreHeatmap(x, annotation_legend = FALSE)
# plotScoreHeatmap(x, labels.use = unique(seurat$cell_type))
# plotScoreHeatmap(x, scores.use = 2, annotation_legend = FALSE)
plotDeltaDistribution(x)
snn_cluster <- "ATAC_snn_res.0.2"
tab <- table(Assigned = seurat[[]][, "cell_type"], Clusters = paste0("G", seurat[[]][, snn_cluster]))
pheatmap(log10(tab + 10), color = colorRampPalette(c('white', brewer.pal(9, "Reds")))(10))

Idents(seurat) <- "cell_type"
idents <- Idents(seurat) %>% 
  unique() %>% 
  na.omit() %>%
  as.character()
cd4 <- subset(
  seurat_integrated, 
  idents = idents %>% .[str_detect(., "CD4+")] 
)

  idents = idents %>% .[str_detect(., "T[_ ]cell")] %>% .[!str_detect(., "CD4")]

  idents = idents %>% .[str_detect(., "B[_ ]cell")]

  idents = idents %>% .[str_detect(., "Monocyte")]
```

```{r, eval = FALSE}
singleR_main <- SingleR(
  test = cts,
  ref = hpca,
  labels = hpca$label.main
)

singleR_fine <- SingleR(
  test = cts,
  ref = hpca2,
  labels = hpca2$label.fine
)

label_type0 <- "label.main"
label_type0 <- "label.fine"

hpca$label_db <- paste0('HPCA.', hpca[[label_type0]])
dice$label_db <- paste0('DICE.', dice[[label_type0]])
# monaco$label_db <- paste0('MONACO.', monaco[[label_type0]])
shared <- intersect(rownames(hpca), rownames(dice))
combined <- cbind(hpca[shared, ], dice[shared, ])
# combined$label_db

# singleR_comb_main <- SingleR(test = cts, ref = combined, labels = combined$label_db)

# ...2. Strategy 2: Comparing scores across references ----------
singleR_comb_fine <- SingleR(
    test = cts,
    ref = list(HPCA = hpca, DICE = dice),
labels = list(hpca[[label_type0]], dice[[label_type0]])
)

singleR_comb3_fine <- SingleR(
  test = cts,
  ref = list(HPCA = hpca, DICE = dice, Monaco = monaco)),
  labels = list(hpca[[label_type0]], dice[[label_type0]], monaco[[label_type0]])
)

# which reference scored best for which label?
grouping <- paste0(singleR_com2$labels,'.', singleR_com2$reference)
best_ref <- as.data.frame(split(singleR_com2, grouping))

# get de. genes from each individual references
metadata(singleR_com2$orig.results$HPCA)$de.genes
metadata(singleR_com2$orig.results$DICE)$de.genes

# ...3. Strategy 3: Using Harmonized Labels ----------

hpca.ont <- celldex::HumanPrimaryCellAtlasData(cell.ont = 'nonna')
dice.ont <- celldex::DatabaseImmuneCellExpressionData(cell.ont = 'nonna')
# monaco.ont <- celldex::MonacoImmuneData(cell.ont = 'nonna')

# Using the same sets of genes:
shared <- intersect(rownames(hpca.ont), rownames(dice.ont))
hpca.ont <- hpca.ont[shared, ]
dice.ont <- dice.ont[shared, ]

# Showing the top 10 most frequent terms:
tail(sort(table(hpca.ont$label.ont)), 10)
tail(sort(table(dice.ont$label.ont)), 10)

# using label.ont instead on label.main while running SingleR
singleR_com3 <- SingleR(
  test = cts,
  ref = list(HPCA = hpca.ont, DICE = dice.ont),
  labels = list(hpca.ont$label.ont, dice.ont$label.ont)
)
```

## Condition markers
```{r feature.heatmaps, eval = FALSE}
FeaturePlot(
  seurat_integrated, 
  features = head(dgea, n = 3) %>% rownames(), 
  split.by = "Patient", 
  min.cutoff = "q10",
  max.cutoff = "q90",
  # cols = c("grey", "red"),
  reduction = 'umap'
)

VlnPlot(
    seurat_integrated, 
    features = head(top_genes, 4) %>% rownames(),
    pt.size = 0, 
    group.by = "cell_type",
    split.by = "Patient", 
    split.plot = TRUE,
    ncol = 2
)

library(ComplexHeatmap)
rowAnnotation(
  "HV" = anno_barplot(top_genes$pct.1),
  "AOSD"= anno_barplot(top_genes$pct.2),
  width = unit(10, "cm")
) %>%
  Heatmap(
    top_genes$avg_log2FC,
    name = "Log2FC",
    cluster_rows = TRUE,
    row_labels = rownames(top_genes),
    right_annotation = .,
    width = unit(1, "cm")
    )

theme_set(theme_cowplot())
pop_seurat <- subset(seurat_integrated, idents = "G1")
Idents(pop_seurat) <- "Patient"
avg_pop <- as.data.frame(log1p(AverageExpression(pop_seurat, verbose = FALSE)$RNA))
avg_pop$gene <- rownames(avg_pop)
p <- ggplot(avg_pop, aes(S01004A, S01013A)) + 
  geom_point() + 
  ggtitle("G1")
LabelPoints(p, points = top_genes %>% head(15) %>% rownames(), repel = TRUE, color = "red")
```

## Pseudo-bulk
```{r scatterplots}
# seurat_integrated@meta.data[, "sample"] <- paste0(
#   seurat_integrated@meta.data[, "Patient"],
#   seurat_integrated@meta.data[, "Type"],
# )
# seurat$cell_type <- seurat[[]][, clusters[nb_optimal]]
# seurat$cell_type <- Idents(seurat) %>% 
  # fct_relevel(as.character(0:9))
cts <- AggregateExpression(
  seurat,
  group.by = c("predicted.labels", "Patient"),
  assay = "RNA",
  slot = "counts",
  return.seurat = FALSE
)

# bulk.mono.de <- FindMarkers(
#   object = cts,
#   ident.1 = "G1_S01004A",
#   ident.2 = "G1_S01013A",
#   test.use = "DESeq2"
# )
# head(bulk.mono.de, n = 15)

cts <- as.data.frame(cts$RNA)
cts <- list.map(
  levels(seurat@meta.data[, "predicted.labels"]),
  f(x) ~{
    select(cts, starts_with(paste0("g", x, "_"))) %>% 
      set_colnames(colnames(.) %>% str_remove_all(paste0("g", x, "_")))
  }
) %>% set_names(levels(seurat@meta.data[, "predicted.labels"]))

saveRDS(cts, file.path(path, "l_cts.rds"))

pct <- pct_by_ident_type(seurat)
saveRDS(pct, file.path(path, "sc_metadata.rds"))
```

```{r}
l_deg0 <- readRDS(file.path(path, "l_deg.rds"))
l_deg <- discard(l_deg0, ~ nrow(.x) == 0)
top10 <- list.map(l_deg,  f(x, xx, xxx) ~head(x, 10) %>% mutate(cluster = xxx)) %>%
    list.rbind() %>%
    rename(gene = "gene_name")
l_bcell_markers <- list.map(l_deg, unique(.) %>% head(12) %>% pull(Gene))
DotPlot(
    seurat, 
    features = unique(pull(top10, "gene")), 
    dot.scale = 8,
    scale = FALSE,
    group.by = "cell_type",
    split.by = "Type",
    cols = GimmeMyPlot:::palette_discrete()
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL) +
    geom_vline(xintercept = last_row_numbers[-last_id] + 0.5) +
    geom_hline(yintercept = seq(2, (length(l_deg0) -1) * 2, by = 2) + 0.5)


l_bcell_markers <- list.map(l_deg, unique(.) %>% pull(Gene))
p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ FeaturePlot(
        seurat,
        features = str_to_sentence(x) %>% .[. %in% Features(seurat)] %>% head(4),
        min.cutoff = "q10", 
        max.cutoff = "q90",
        cols = c("white", "red"),
        split.by = "Type",
        reduction = "umap",
        pt.size = 1,
        alpha = .5,
        combine = FALSE
    )  %>% 
  map(~theme_sc_dim(.) + labs(x = NULL, y = NULL)) %>%
  plot_grid(plotlist = ., ncol = 4) +
    plot_annotation(
      title = z, 
      theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
    ) 
)

p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ 
VlnPlot(
    object = seurat,
    features = str_to_sentence(x) %>% .[. %in% Features(seurat)] %>% head(12),
    pt.size = 0,
    group.by = "cell_type",
    split.by = "Type",
    cols = cols,
    combine = FALSE
) %>%
map(~. +theme_violin() + labs(y = NULL) + NoLegend()) %>%
  plot_grid(plotlist = ., ncol = 4, nrow = 3) +
    plot_annotation(
      title = z, 
      theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
    ) 
)
# i = 2; plot_grid(plotlist = p [(12*(i-1)+1):(12*i)], ncol = 4, nrow = 3)


list.map(
    split(cell_markers_bcell, ceiling(seq_along(cell_markers_bcell)/4)),
f(x) ~ 
FeaturePlot(
    seurat,
    features = x,
    min.cutoff = "q10", 
    max.cutoff = "q90",
    cols = c("white", "red"),
    split.by = "Type",
    reduction = "umap",
    pt.size = 1,
    alpha = .5,
    combine = FALSE
)  %>% 
    map(~theme_sc_dim(.) + labs(x = NULL, y = NULL))  %>%
    plot_grid(plotlist = ., ncol = 4)
)

top10 <- l_bcell_markers %>%
  list.map(f(x, y, z) ~  x[x %in% Features(seurat)] %>% head(10) %>% data.frame(gene = ., cluster = y)) %>% 
  list.rbind() %>%
  filter(!duplicated(gene))

DotPlot(
    seurat, 
    features = list.map(l_bcell_markers, .[. %in% Features(seurat)] %>% head(10)) %>% unlist() %>% unique(), 
    dot.scale = 8,
    scale = FALSE,
    group.by = "cell_type",
    split.by = "Type",
    cols = GimmeMyPlot:::palette_discrete()
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL) +
    geom_vline(xintercept = last_row_numbers[-last_id] + 0.5) +
    # geom_vline(xintercept = seq(10, 60, 10) + 0.5) +
    geom_hline(yintercept = seq(2, 12, by = 2) + 0.5)

list.map(
    top50,
    f(i, ii, z) ~
    list.map(
        str_to_sentence(i) %>% .[. %in% Features(seurat)] %>% head(4*3) %>% split(., ceiling(seq_along(.)/4)),
    f(x) ~ 
        FeaturePlot(
            seurat,
            features = x,
            min.cutoff = "q10", 
            max.cutoff = "q90",
            cols = c("white", "red"),
            split.by = "Type",
            reduction = "umap",
            pt.size = 1,
            alpha = .5,
            combine = FALSE
        )  %>% 
        map(~theme_sc_dim(.) + labs(x = NULL, y = NULL))  %>%
        plot_grid(plotlist = ., ncol = 4) +
        plot_annotation(
            title = z, 
            theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
        )
)
) -> p
```

```{r }
cts0 <- cts[[1]]
colData <- data.frame(samples = colnames(cts0)) %>%
  mutate(condition = str_extract(samples, unique(seurat$Type) %>% paste(collapse = "|")) %>% as.factor()) %>%
  column_to_rownames(var = 'samples')

dds <- DESeqDataSetFromMatrix(
  countData = cts0,
  colData = colData,
  design = ~ condition
) %>%
   .[rowSums(counts(.) >= 10) >= 3, ] %>%
  DESeq(quiet = TRUE)


resultsNames(dds)[2]
res <- results(dds, name = resultsNames(dds)[2])

# genes.to.label <- NULL
# p <- lapply(
#   c("S01004A", "S01013A"),
#   function(x) {
#   CellScatter(aggregate_seurat, paste0("G0_", x), paste0("G1_", x), highlight = genes.to.label) %>%
#   LabelPoints(points = genes.to.label, repel = TRUE)
# })
```

```{r}
colours <- get_colors()[c(3, 1, 4:5)]
list.map(tmpp0, f(x, y) ~data.frame(Type = colours[y], val = x)) %>% list.rbind() %>% set_rownames(rownames(.) %>% str_remove_all("A\\d{2}PB\\d")) -> cou
plot_bar(select(cou, 2), colour = pull(cou, 1), label_x = "none", ratio = 10) + theme(axis.text.y = element_text(colour = "gray"), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), panel.grid.major.y = element_line(colour = "grey", linewidth = 0.5), panel.grid.major.x = element_blank()) + ggtitle(NULL)
```

# Trajectory analysis

## Monocle
```{r}
library(monocle3)
library(SeuratWrappers)
cds <- as.cell_data_set(seurat)
```

```{r, eval = FALSE}
expression_data <- GetAssayData(seurat, assay = "RNA", slot = "counts")
gene_annotation <- data.frame(
  gene_short_name = rownames(expression_data),
  row.names = rownames(expression_data)
)

cds <- new_cell_data_set(
  expression_data,
  cell_metadata = seurat@meta.data,
  gene_metadata = gene_annotation
)

reducedDims(cds)$UMAP <- Embeddings(seurat, reduction = "umap")
cds@clusters$UMAP$clusters <- seurat@active.ident

partitions <- factor(rep(1, ncol(cds)))
cds@clusters$UMAP$partitions <- partitions
```

```{r}
names(cds@clusters$UMAP$partitions) <- colnames(cds)
root_cells <- colnames(cds)[cds@clusters$UMAP$clusters == "5"]

cds <- learn_graph(cds, use_partition = FALSE) %>%
  order_cells(root_cells = root_cells)

cds@principal_graph_aux@listData$UMAP$pseudotime <- max(pseudotime(cds), na.rm = TRUE) - pseudotime(cds)
```

```{r}
plot_cells(
    cds,
    color_cells_by = "pseudotime",
    label_root = FALSE,
    label_leaves = FALSE,
    label_branch_points = FALSE,
    trajectory_graph_color = "black",
    trajectory_graph_segment_size = 1.5,
    cell_size = 1,
    alpha = .15
) %>% theme_sc_dim() + 
    scale_color_gradientn(
      colors = col_pal, 
      name = "Pseudotime"
    )

# seurat <- as.Seurat(cds, assay = "integrated")
```

## Slingshot
```{r}
library("slingshot")
start.clus <- 5

sce <- as.SingleCellExperiment(seurat)
reducedDims(sce)$UMAP <- Embeddings(seurat, reduction = "umap") -> reduc_coords

sce <- slingshot(
  sce,
  start.clus = start.clus,
  clusterLabels = "cell_type",
  reducedDim = "UMAP"
)

# lin <- getLineages(reduc_coords, seurat$cell_type, start.clus = start.clus)

# SlingshotDataSet(sce)

curves <- slingCurves(sce)
lineage_probs <- list.map(curves, .$w) %>%
  list.cbind() %>%
  max.col(ties.method = "first")

pseudotime <- slingPseudotime(sce) %>%
  .[cbind(seq_len(nrow(.)), lineage_probs)]

plot(
  reduc_coords,
  # col = GimmeMyPlot:::palette_discrete()[sce$cell_type],
  col = colorRampPalette(col_pal)(100)[cut(pseudotime, breaks = 100)],
  pch = 16,
  asp = 1
)
# lines(SlingshotDataSet(sce), lwd = 4, col = "black")
lines(SlingshotDataSet(sce), lwd = 4, type = 'lineages', col = "black",  show.constraints = TRUE)

p <- as.data.frame(reduc_coords) %>%
  mutate(pseudotime = pseudotime) %>%
  ggplot(aes(x = umap_1, y = umap_2, colour = pseudotime)) +
  geom_point(alpha = 0.8)

p <- reduce(
  names(curves),
  .f = function(plot, lineage_name) {
    plot + geom_path(
      data = as.data.frame(curves[[lineage_name]]$s),
      aes(x = umap_1, y = umap_2),
      size = 1.2,
      color = "black",
      inherit.aes = FALSE
    )
  },
  .init = p
)
theme_sc_dim(p)  + 
    scale_color_gradientn(
      colors = col_pal, 
      name = "Pseudotime"
    )
```

# Transfer learning
```{r}
wt@meta.data[, "labels"] <- Idents(wt)
# wt@assays$SCT@scale.data <- matrix()
# ko@assays$SCT@scale.data <- matrix()
seurat <- merge(wt, y = ko)
```

```{r}
anchors <- FindTransferAnchors(
    reference = wt,
    query = seurat,
    # normalization.method = "SCT",
    reference.reduction = "pca",
    # reference.reduction = integration_method,
    dims = seq(n_dim)
)
```

```{r, eval = FALSE}
cl_trans <- TransferData(
    anchorset = anchors,
    dims = seq(n_dim),
    # weight.reduction = KO[[integration_method]],
    refdata = Idents(wt)
  )
seurat <- AddMetaData(seurat, metadata = cl_trans)
```

```{r, eval = FALSE}
wt2 <- RunUMAP(
  wt,
  dims = seq(n_dim),
  # reduction = integration_method,
  reduction = "pca",
  return.model = TRUE
)
```

```{r}
seurat <- MapQuery(
    anchorset = anchors,
    query = seurat,
    reference = wt,
    refdata = list(labels = "labels"),
    # reference.reduction = integration_method,
    reduction.model = "umap"
)
```

```{r}
plot_sc_violin(
    seurat,
    features = paste0("prediction.score.", levels(Idents(wt))),
    cols = GimmeMyPlot:::palette_discrete(),
    nrow = 3
)
```

```{r}
type <- NULL
seurat$predicted.labels<- factor(seurat$predicted.labels, levels = 0: (length(unique(seurat$predicted.labels)) - 1))
Idents(seurat) <- "predicted.labels"
DimPlot(
    object = seurat, 
    label = TRUE,
    # cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    cols = GimmeMyPlot:::palette_discrete(),
    repel = TRUE,
    na.value = "white",
    split.by = "Type",
    # group.by = ,
    reduction = "ref.umap",
    pt.size = 1,
    alpha = 0.25
) %>% 
    theme_sc_dim() + ggtitle(type) + NoLegend()
```

## Session information
```{r end, echo = FALSE}
sessionInfo()
```
