---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup
```{r init, message = FALSE}
library(GimmmeMySC)
library(RColorBrewer)
library(rlist)
library(Seurat)
library(cowplot)
library(GimmeMyPlot)
library(scDblFinder)

set.seed(1234)
options(future.globals.maxSize = 2000 * 1024^2)
options(ggrepel.max.overlaps = 1000)

path <- file.path(
  dirname(getwd()),
  "inst",
  "extdata"
)

SCT <- TRUE
max_dim <- 100
n_dim <- 30
assay <- "RNA"
species <- "human"

l_samples <- list.files(path) %>%  
  .[sapply(., function(x) dir.exists(file.path(path, x)))]  %>%
  .[str_detect(., "KO")]
col_inds0 <- palette_discrete()[seq_along(l_samples)]

if (species == "human") {
    func_format <- identity
} else {
    func_format <- str_to_sentence
}
```

# Load data
```{r data}
list_seurat <-  list.map(
    l_samples,
    f(x) ~{
      count_matrix <- file.path(path, x, "outs") %>%
        list.files(full.names =TRUE) %>% 
        .[str_detect(., "filtered_feature_bc_matrix.h5")]
      tmp <- Read10X_h5(count_matrix) %>%
      CreateSeuratObject()
    tmp[["percent_mitochondrial"]] <- PercentageFeatureSet(tmp, pattern = "^((MT)|(mt))-")
    tmp
    }
)
```

```{r data}
types  <- str_remove_all(l_samples, "_?\\d+")

col_inds <- list(
  c("#FB9A99", "#E31A1C"),
  c("#B2DF8A", "#33A02C")
)
color_types <- sapply(col_inds, last)

col_inds <- list.mapv(unique(types), f(i, j) ~colorRampPalette(col_inds[[j]])(length(types[types == i]))) %>% as.character()

seurat <- merge(
  x = list_seurat[[1]],
  y = list_seurat[-1],
  add.cell.ids = paste0(
      types,
      "_",
      # str_remove_all(l_samples, types) %>% str_remove_all("_")
      l_samples
    )
)

seurat$sample <- rownames(seurat[[]])
seurat[[]] <- separate(
  seurat[[]], 
  col = "sample", 
  into = c("Type", "Patient", "Barcode"),
  sep = "_"
)
```

# QC
## Filter cells with low or too high genes
```{r qc_plot}
print_sc_stats(seurat, probs = c(0, 0.1, 0.5, 0.9, 1)) %>%
  kable_sc(digits = 3)

# plot_sc_cor(seurat)

list.map(
  c(paste0("nFeature", "_", assay), "percent_mitochondrial"),
  f(i, j) ~ {
    density_scatter(
      seurat,
      x = i,
      y = paste0("nCount", "_", assay),
      log_x = TRUE,
      log_y = TRUE
  )
  }
    
) %>% plot_grid(plotlist = ., align = "hv")

Idents(seurat) <- "Patient"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10)
)
```

```{r}
seurat_old <- seurat
list_seurat0 <- list_seurat
features <- c("nFeature_RNA", "nCount_RNA", "percent_mitochondrial")
list_seurat <- list.map(
    list_seurat,
    f(i) ~{
      stats_qc <- sapply(
        features,
        function(x) quantile(
            i[[x]], 
            probs = c(0, .025, .05, .1, seq(.25, .75, .25), .9, .95, .975, 1), 
            na.rm = TRUE)
    )
    subset(
      i, 
      subset = 
        nFeature_RNA > stats_qc["10%", "nFeature_RNA"] &
        nFeature_RNA < stats_qc["90%", "nFeature_RNA"] & 
        percent_mitochondrial < stats_qc["90%", "percent_mitochondrial"] & 
        nCount_RNA < stats_qc["90%", "nCount_RNA"]
  )
  }
)

list.mapv(
  seq_along(list_seurat0),
  f(i) ~ print_filtered_cells(list_seurat0[[i]], list_seurat[[i]])
)

seurat <- integrate_multisamples(list_seurat, paste0(types, "_", l_samples))
```

## Doublets
```{r}
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
sce <- scDblFinder(
  GetAssayData(seurat[["RNA"]], slot = "counts"),
  samples = seurat@active.ident,
  aggregateFeatures = FALSE,
  k = NULL,
  processing = "normFeatures",
  dims = 20
)
table(sce$scDblFinder.class)

seurat$doublets <- sce[, row.names(seurat[[]])]$scDblFinder.class

df_doublets <- list.map(
    unique(seurat$Patient),
    f(x) ~ filter(seurat[[]], Patient == x) %>% pull(doublets) %>% table()
) %>% 
    list.rbind() %>%
    as.data.frame() %>%
    rbind(".Total" = colSums(.))

t(df_doublets) %>% 
  plot_bar_2cat(
    count = TRUE,
    stats = FALSE,
    colour = brewer.pal(3, "Set1")[c(1, 3)],
    colour_text = "black"
  )
```

```{r}
# visualize doublets
# DimPlot(
#   seurat, 
#   reduction = 'umap.unintegrated', 
#   group.by = "doubletFinder",
#   label = TRUE,
#   repel = TRUE
# )

# number of singlets and doublets

Idents(seurat) <- "doublets"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10),
    cols = palette_discrete()
)
Idents(seurat) <- "Patient"

seurat <- subset(seurat, subset = doublets == "singlet")
```

## QC after
```{r}
print_sc_stats(seurat, probs = c(0, 0.1, 0.5, 0.9, 1)) %>%
  kable_sc(digits = 3)

list.map(
  c(paste0("nFeature", "_", assay), "percent_mitochondrial"),
  f(i, j) ~ {
    density_scatter(
      seurat,
      x = i,
      y = paste0("nCount", "_", assay),
      log_x = TRUE,
      log_y = TRUE
  )
  }
    
) %>% plot_grid(plotlist = ., align = "hv")

Idents(seurat) <- "Patient"
plot_sc_violin(
    seurat,
    cols = col_inds,
    nrow = 1,
    normalize = c(10, 10, 10)
)
```

```{r qc_sub}
(ncol(seurat) / ncol(seurat_old) * 100 - 100) %>% `*`(-1) %>% round(1) %>% paste("%")

list.map(
    l_samples,
    f(x) ~ list.map(
      c(seurat_old, seurat),
      f(i) ~ filter(i[[]], Patient == x) %>% nrow()
      ) %>% 
    list.rbind()
) %>% 
    list.cbind() %>%
    as.data.frame() %>% 
    set_colnames(l_samples) %>%
    t() %>%
    as.data.frame() %>%
    rbind(`.Total` = colSums(.)) %>%
    mutate(V3 = V1 - V2) %>%
    select(-V1) %>%
    t() %>%
    as.data.frame() %>%
    set_rownames(c("Kept", "Removed"))  %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE
    )
```

## Negative / positive control genes
```{r}
g_neg <- c("DRD1", "ALB", "MYL1", "NPPA") %>% sort()
g_pos <- c("GAPDH", "ACTB", "TMSB4X", "B2M") %>% sort()
g_ctr <- c(g_neg, g_pos)
g_sex <- c("XIST", "DDX3Y")
list.map(
    list(g_sex, g_ctr),
    ~DotPlot(
    seurat, 
    features = func_format(.), 
    dot.scale = 8
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL)
)%>%
    plot_grid(plotlist = ., ncol = 2, align = "hv")
```

# Gene filtering
```{r filter_mark}
if (!SCT)
    variable_plot(x)
```

# PCA
```{r pca}
s.genes <- cc.genes$s.genes %>% func_format()
g2m.genes <- cc.genes$g2m.genes %>% func_format()

if (!SCT) {
  if (assay == "sketch") {
    seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
  seurat <- ScaleData(seurat)
} else {
  seurat[["RNA"]] <- split(
    seurat[["RNA"]], 
    f = seurat$Patient,
    layers = c("counts","data")
  )
  seurat <- SCTransform(seurat)
  assay <- "SCT"
  DefaultAssay(seurat) <- "SCT"
  if (assay == "sketch") {
     seurat <- SketchData(seurat, over.write = TRUE)
    DefaultAssay(seurat) <- "sketch"
  }
}
seurat <- RunPCA(seurat, reduction_key = "Dim. ", npcs = max_dim)

# Dimension visualisation
DimPlot(
    seurat, 
    # cells = seq(3.2e5),
    reduction = 'pca',
    group.by = "Patient", 
    split.by = "Type", 
    cols = col_inds,
    repel = TRUE,
    pt.size = .5,
    alpha = .25
) %>% theme_sc_dim()
DimPlot(
  seurat, 
  reduction = 'pca',
  group.by = "Type", 
  cols = color_types
) %>% theme_sc_dim()
# coord <- trans_coord(
#   seurat@reductions$pca@cell.embeddings[, seq(2)], 
#   add = seurat@meta.data[, "seurat_clusters"] %>% paste("Cluster", .)
# )
# plot_dimred(coord, eig, colour = GimmeMyPlot:::palette_discrete2())
```

## Dimension selection
```{r}
# Dimension selection
plot_eig(seurat[['pca']]@stdev) + 
   geom_hline(yintercept = mean(seurat[['pca']]@stdev), col = "gray50", lwd =1, lty = 2)

eig <- seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100
p1 <- plot_eig(eig) + 
  geom_hline(yintercept = mean(eig), col = "gray50", lwd = 1, lty = 2) +
  ylab("% Explained\n variance")

p2 <- cumsum(eig) %>% 
  plot_eig()+ 
  ylim(0, NA) +
  ylab("% Cumulated\n explained variance") + 
  geom_hline(yintercept = 80, color = "gray50", lwd =1, lty = 2)

p1 + p2

cumsum(eig) %>% .[. < 80] %>% last()
```

```{r}
DimHeatmap(seurat, dims = seq(15), cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(seurat, dims = 31:45, cells = 500, balanced = TRUE)
```

```{r}
DimPlot(
  seurat,
  reduction = "pca",
  group.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  pt.size = .5,
  alpha = .25
) %>% theme_sc_dim() + ggtitle(NULL)

DimPlot(
  seurat,
  reduction = "pca",
  split.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  pt.size = .5,
  alpha = .25
) %>% theme_sc_dim() +
    ggtitle(NULL) + 
    NoLegend()
```

```{r}
comp_genes <- list.map(
    seq(max_dim),
    f(x) ~ Loadings(seurat, reduction = "pca") %>%
        as.data.frame() %>%
        select(paste0("PC_", x)) %>%
        arrange(abs(!!sym(paste0("PC_", x)))) %>%
        tail(20) %>%
        rownames()
) %>% set_names(seq(max_dim))

list.map(
  list(g2m.genes, s.genes),
  f(i) ~lapply(comp_genes, function(x) x[x %in% func_format(i)]) %>% compact()
) %>% set_names(c("G2/M", "S"))

# n_dim <- 20
round(cumsum(eig)[n_dim], 1)
```


## Component markers
```{r pca_plot}
# DimHeatmap(seurat, dims = 1)

# Markers
print(seurat[['pca']], dims = 1:5, nfeatures = 5)
VizDimLoadings(seurat, dims = 1:2, reduction = 'pca')

# i <- 1
# seurat[['pca']]@feature.loadings[, i] %>%
#   plot_bar(
#     abs = TRUE,
#     n_max = 10,
#     label_x = "none",
#     geom = "point",
#     title = paste0("Dim ", i, " (", round(seurat[['pca']]@stdev / sum(seurat[['pca']]@stdev) * 100, 1) , "%)"),
#     colour = "red"
#   )
```


## Batch effect
```{r umap}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction.key = "Dim. ",
  )
```

```{r umap}
DimPlot(
  seurat,
  group.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  pt.size = .5,
  alpha = .25
) %>% theme_sc_dim() + ggtitle(NULL)
DimPlot(
  seurat,
  group.by = 'Patient',
  split.by = 'Patient',
  cols = col_inds0,
  repel = TRUE,
  alpha = .25
) %>% theme_sc_dim() + NoLegend()
```


```{r, eval = FALSE}
DimPlot(
  seurat,
  reduction = 'umap.unintegrated',
  group.by = 'Type',
  cols = color_types,
  repel = TRUE
) %>% theme_sc_dim() +
DimPlot(
  seurat,
  reduction = 'umap.unintegrated',
  group.by = 'seurat_clusters',
  cols = col_clus,
  label = TRUE,
  repel = TRUE
)

DimPlot(
  seurat,
  # reduction = 'umap.unintegrated',
  group.by = "Patient",
  split.by = "Type",
  cols = col_inds,
  repel = TRUE
) %>% theme_sc_dim()
```

# Integration
```{r, eval = FALSE}
if(SCT) {
  seurat[["v5"]] <- CreateAssay5Object(
    counts =  seurat@assays$RNA$counts,
    data =  seurat@assays$SCT$data,
    # scale.data = seurat@assays$SCT$scale.data
  )
  seurat@assays[["v5"]]@key <- "v5_"
  DefaultAssay(seurat) <- "v5"
    seurat[["v5"]] <- split(
    seurat[["v5"]], 
    f = seurat$Patient,
    layers = c("counts","data")
  )
  assay <- "V5"
}
```

```{r}
integration_method <- "RPCA"
seurat <- IntegrateLayers(
    object = seurat, 
    method = get(paste0(integration_method, "Integration")),
    new.reduction = integration_method,
    # reduction.key = "Dim. ",
    orig.reduction = "pca",
    assay = assay,
    dims.to.integrate = seq(n_dim),
    normalization.method = ifelse(SCT, "SCT", "LogNormalize"),
    verbose = TRUE
)

seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
```

## UMAP
```{r}
seurat <- RunUMAP(
  seurat, 
  dims = seq(n_dim),
  reduction = integration_method
)
```

```{r int}
assay_old <- assay
assay <- "RNA"
p <- map(
    features,
    ~FeaturePlot(
        object = seurat,
        features = .,
        alpha = 0.5,
        # cols = c("yellow", "blue"), 
        # min.cutoff = "q10", 
        # max.cutoff = "q90"
    ) %>% 
        theme_sc_dim() +
        ggtitle(str_clean(.)) +
    scale_color_gradientn(
        colors = pal_sc
    )

)

plot_grid(plotlist = p, nrow = 2, ncol = 2, align = "hv")
assay <- assay_old
```


```{r}
list.map(
  c("pca", integration_method),
  f(x) ~{
  DimPlot(
    seurat, 
    reduction = x, 
    group = "Patient",
    cols = col_inds0,
    label = TRUE,
    repel = TRUE
) %>% theme_sc_dim() + ggtitle(x)
  }
) %>% plot_grid(plotlist = ., ncol = 2)
```

## Cell cycle
```{r}
seurat <- CellCycleScoring(
  seurat,
  s.features = s.genes, 
  g2m.features = g2m.genes, 
  set.ident = TRUE
)
seurat$CC.Difference <- seurat$S.Score - seurat$G2M.Score

p1 <- FeaturePlot(
    seurat,
    features = "CC.Difference",
    alpha = 0.5
) %>% theme_sc_dim()  +
    scale_color_gradientn(
        colors = pal_sc,
        limits = c(-1, 1),
        name = "S to G2/M"
    ) + labs(x = NULL, y = NULL, title = NULL)


p2 <- DimPlot(
    object = seurat, 
    label = TRUE,
    # cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    cols = GimmeMyPlot:::palette_discrete(),
    repel = TRUE,
    group.by = "Phase",
    na.value = "white",
    reduction = "umap",
    pt.size = 1,
    alpha = 0.25
) %>% 
    theme_sc_dim() +  labs(x = NULL, y = NULL, title = NULL)

p1 + p2

RidgePlot(seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67") %>% func_format(), ncol = 2)
```

## Automatic annotation
```{r}
if (species == "human") {
    hpca <- celldex::HumanPrimaryCellAtlasData()
    dice <- celldex::DatabaseImmuneCellExpressionData()
} else {
    hpca <- celldex::ImmGenData()
    dice <- celldex::MouseRNAseqData()
}

hpca$label.main <- format_annot(hpca$label.main)
dice$label.main <- format_annot(dice$label.main)
hpca$label.fine <- format_annot(hpca$label.fine)
dice$label.fine <- format_annot(dice$label.fine)

bcell_subtypes <- unique(hpca$label.fine) %>%
  .[str_detect(., "FR[A-F]|CLP")] %>%
  sort() %>%
  rev()
is_bcell <- str_starts(hpca$label.fine, "B cell")
keep_cells <- (!is_bcell) | (hpca$label.fine %in% bcell_subtypes)

hpca2 <- hpca[, keep_cells]

cts <- GetAssayData(seurat, layer = 'counts')
```

```{r}
singleR_main <- SingleR(
  test = cts,
  ref = hpca,
  labels = hpca$label.main
)

singleR_fine <- SingleR(
  test = cts,
  ref = hpca2,
  labels = hpca2$label.fine
)

annotation <- list(singleR_main = singleR_main, singleR_fine = singleR_fine)
saveRDS(annotation, file = file.path(path, paste(project,  "_annotation.rds")))
```

```{r, eval = FALSE}
label_type0 <- "label.main"
label_type0 <- "label.fine"

hpca$label_db <- paste0('HPCA.', hpca[[label_type0]])
dice$label_db <- paste0('DICE.', dice[[label_type0]])
shared <- intersect(rownames(hpca), rownames(dice))
combined <- cbind(hpca[shared, ], dice[shared, ])


# ...2. Strategy 2: Comparing scores across references ----------
singleR_comb <- SingleR(
    test = cts,
    ref = list(HPCA = hpca, DICE = dice),
labels = list(hpca[[label_type0]], dice[[label_type0]])
)

# which reference scored best for which label?
grouping <- paste0(singleR_comb$labels,'.', singleR_comb$reference)
best_ref <- as.data.frame(split(singleR_comb, grouping))

# get de. genes from each individual references
metadata(singleR_comb$orig.results$HPCA)$de.genes
metadata(singleR_comb$orig.results$DICE)$de.genes

# ...3. Strategy 3: Using Harmonized Labels ----------

hpca.ont <- HumanPrimaryCellAtlasData(cell.ont = 'nonna')
dice.ont <- DatabaseImmuneCellExpressionData(cell.ont = 'nonna')

# Using the same sets of genes:
shared <- intersect(rownames(hpca.ont), rownames(dice.ont))
hpca.ont <- hpca.ont[shared, ]
dice.ont <- dice.ont[shared, ]

# Showing the top 10 most frequent terms:
tail(sort(table(hpca.ont$label.ont)), 10)
tail(sort(table(dice.ont$label.ont)), 10)

# using label.ont instead on label.main while running SingleR
singleR_ont <- SingleR(
  test = cts,
  ref = list(HPCA = hpca.ont, DICE = dice.ont),
  labels = list(hpca.ont$label.ont, dice.ont$label.ont)
)
```

```{r}
label_type <- "labels"
x1 <- "singleR_fine"
x <- get(x1)
# x$pruned.labels
# x$scores
x0 <- as.data.frame(x) %>% select(all_of(label_type))
seurat[[]][, x1] <- x0

if (pull(x0, label_type) %>% str_detect("CL\\:\\d*") %>% any()) {
  x0 <- rbind(as.data.frame(colData(hpca.ont)), as.data.frame(colData(dice.ont))) %>%
      filter(str_detect(label.ont, paste0("^", unique(pull(x0, "labels")), "$", collapse = "|"))) %>% 
      filter(!duplicated(label.ont)) %>%
      mutate(labels = label.ont) %>%
      mutate(label.main = label.main %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
    mutate(label.main = label.fine %>% 
        str_replace_all("(cell)s", "\\1") %>%
        str_replace_all("_", " ") %>%
        str_replace_all("\\:", ", ") %>%
        str_replace_all("(Monocyte)s", "\\1")) %>%
      left_join(x0) %>%
      column_to_rownames()
  label_type <- "label.fine"
}

to_remove <- table(pull(x0, label_type)) %>%
    .[. < lim] %>%
    names()
x0[pull(x0, label_type) %in% to_remove, label_type] <- NA
pull(x0, label_type) %>% 
    cluster_size() %>% 
    kable_sc()
pull(x0, label_type) %>%
    cluster_size() %>% 
  select(n) %>%
  GimmeMyPlot::plot_bar(
    label_x = "none",
    label_y = "percent",
    sample_size = nrow(x0),
    digits = 1,
    n_max = 15
  ) + ggtitle(NULL)
if (x1 == "singleR_main") {
seurat$cell_type <- 
  x0[match(rownames(seurat[[]]), rownames(x0)), label_type]
}

to_keep <- paste(paste(unique(seurat$cell_type) %>% na.omit(),  "\\("), collapse = "|")
x0[[label_type]][!str_detect(x0[[label_type]], to_keep)] <- NA

seurat$b_cell <- seurat$cell_subtype
seurat$b_cell[!str_detect(seurat$b_cell, "B cell")] <- NA
seurat$b_cell <- keep_parenthesis(seurat$b_cell)

x0 <- data.frame(labels = seurat$singleR_main)

seurat$cell_type_formatted <- remove_parenthesis(seurat$cell_subtype)
seurat$cell_subtype_formatted <- keep_parenthesis(seurat$cell_subtype)
```

```{r}
DimPlot(
    seurat, 
    reduction = 'umap', 
    group.by = "cell_type",
    cols = palette_discrete(),
    label = TRUE,
    repel = TRUE,
    alpha = 0.5,
    na.value = "white"
) %>% theme_sc_dim() + ggtitle(NULL)

DimPlot(
    seurat, 
    reduction = 'umap', 
    split.by = "cell_type",
    group.by = "cell_type",
    cols = palette_discrete(),
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white",
    ncol = 3
) %>% theme_sc_dim() + ggtitle(NULL)
```

```{r}
DimPlot(
    seurat, 
    reduction = 'umap', 
    group.by = "cell_subtype",
    cols = palette_continuous(gray = FALSE)(unique(seurat$cell_subtype) %>% length()),
    label = TRUE,
    repel = TRUE,
    alpha = 0.5,
    na.value = "white"
) %>% theme_sc_dim() + ggtitle(NULL)

list.map(
    unique(seurat$cell_type_formatted) %>% na.omit(),
    f(x) ~subset(seurat, cell_type_formatted == x)
    ) %>%
list.map(
    f(x, y, z) ~
        DimPlot(
            x, 
            reduction = 'umap',
            group.by = "cell_subtype_formatted",
            na.value = "white",
            cols = rep(palette_discrete(), 5),
            label = TRUE,
            repel = TRUE
        )  %>% theme_sc_dim() +
        labs(title = z, x = NULL, y = NULL) 
) %>% plot_grid(plotlist = ., align = "hv", nrow = 2)

cell_categories <- unique(seurat$cell_type_formatted)
cell_list <- map(cell_categories, ~ {
  seurat$cell_subtype[str_detect(seurat$cell_subtype, paste0("^", .x))] %>% na.omit() %>% keep_parenthesis()
}) %>%
  set_names(cell_categories)

p <- list.map(
  compact(cell_list),
  f(x, y, z) ~ cluster_size(x) %>% 
    select(n) %>%
    GimmeMyPlot::plot_bar(
        label_x = "none",
        label_y = "percent",
        sample_size = length(x),
        digits = 1,
        n_max = 15,
        ratio = 1
    ) + ggtitle(z)
)

plot_grid(plotlist = p, align = "hv", ncol = 3, nrow = 2)
```

```{r}
zoom_coord <- function(x) x + xlim(-7, 5) + ylim(-7, 8)
DimPlot(
    seurat, 
    reduction = 'umap', 
    group.by = "b_cell",
    cols = palette_discrete(),
    label = TRUE,
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white"
) %>% theme_sc_dim() + ggtitle(NULL) %>%
    zoom_coord()

DimPlot(
    seurat, 
    reduction = 'umap', 
    split.by = "b_cell",
    group.by = "b_cell",
    cols = palette_discrete(),
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white",
    ncol = 3
) %>% theme_sc_dim() + ggtitle(NULL) %>%
    zoom_coord()
```

```{r}
plotScoreHeatmap(x, labels.use = unique(seurat$cell_subtype), annotation_legend = FALSE)
# plotScoreHeatmap(x, scores.use = 2, annotation_legend = FALSE)
plotDeltaDistribution(x)
```


## Cluster
```{r clus_int}
resolution2 <- seq(.1, .4, 0.05)
resolution <- c(seq(.5, 1, by = 0.1), 1.2, 1.5)
resolution_tot <-c(resolution2, resolution)
DefaultAssay(seurat) <- assay
seurat <- FindNeighbors(
  seurat,
  reduction = integration_method, 
  dims = seq(n_dim)
)

seurat <- FindClusters(
    seurat,
    verbose = FALSE,
    algorithm = 3,
    resolution = resolution_tot,
    graph.name = tail(names(seurat@graphs), 1)
  )

seurat2 <- seurat
prefix_cl <- paste0(assay, "_snn_res.")
cols_to_remove <- paste0(prefix_cl, resolution[-1])
seurat2@meta.data <- seurat2@meta.data[, !colnames(seurat2@meta.data) %in% cols_to_remove]

select(seurat@meta.data, contains(prefix_cl)) %>%
     list.mapv(f(x) ~ levels(x) %>% length()) %>% 
    set_names(names(.) %>% str_remove(prefix_cl)) %>%
  plot_bar(
    label_x = "none",
    label_y = "value",
    sort = "desc",
    digits = 1,
    n_max = 15
  ) + ggtitle(NULL)
```
```{r}
clusters <- grep(prefix_cl, colnames(seurat[[]]), value = TRUE) %>% sort()
clusters2 <- paste0(prefix_cl, resolution2)
list.mapv(clusters, f(i) ~length(unique(seurat@meta.data[, i])))
cl_size <- map_dfr(
    clusters,
    ~ cluster_size(seurat@meta.data[, .x]) %>% 
        pull(1) %>% 
        tibble(
            Clusters = .x,
            cluster = seq_along(.),
            inertia = .
        )
) %>% pivot_wider(names_from = cluster, values_from = inertia, names_prefix = "") %>%
  as.data.frame() %>%
  mutate(Clusters = str_remove_all(Clusters, paste0(assay, "_snn_res.")))
  
cl_size[is.na(cl_size)] <- ""
kable0(cl_size)
```

```{r}
clustree(seurat, prefix = prefix_cl) +
    # scale_color_discrete(type = brewer.pal(length(clusters), "Reds")) +
    scale_color_discrete(type = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(clusters))) +
    theme(
        legend.title = element_text(face = "italic", size = 15),
        legend.text = element_text(size = 10)
    )
```

```{r}
p_cl <- list.map(
    clusters,
    f(x, y, z) ~
        DimPlot(
            seurat, 
            reduction = 'umap',
            group.by = x,
            na.value = "white",
            cols = palette_continuous()(seurat@meta.data[, x] %>% unique() %>% length()),
            label = TRUE,
            repel = TRUE
        )  %>% theme_sc_dim() + 
      NoLegend() + 
      labs(x = NULL, y = NULL, title = str_remove(x, ".*_snn_res."))
)
plot_grid(plotlist = p_cl[seq(8)], align = "hv", ncol = 4, nrow = 2)
plot_grid(plotlist = p_cl[8:15], align = "hv", ncol = 4, nrow = 2)
```


```{r clus_int_res}
best_resolution <- 0.35
snn_cluster <- paste0(
  ifelse(SCT, "SCT", assay),
  "_snn_res.",
  best_resolution
)

Idents(seurat) <- seurat[[]][, snn_cluster] #%>%
  # as.character() %>% 
  # as.numeric() %>% 
  # `+`(1) %>%
  # str_replace_all("^(\\d)$", "0\\1") %>%
  # paste("G", .)

Idents(seurat) %>% 
    cluster_size() %>%
    kable_sc()

# names(cell_type) <- levels(seurat_integrated)
# seurat_integrated <- RenameIdents(seurat_integrated, cell_type)
```


```{r}
DimPlot(
  seurat, 
  reduction = "umap",
  cols = palette_continuous()(length(levels(Idents(seurat)))),
  label = TRUE,
  repel = TRUE
  )

DimPlot(
    seurat, 
    reduction = 'umap', 
    split.by = snn_cluster,
    cols = palette_continuous()(length(levels(Idents(seurat)))),
    repel = TRUE,
    pt.size = 0.5,
    na.value = "white",
    ncol = 4
) %>% theme_sc_dim() + 
    ggtitle(NULL)

table(Idents(seurat), seurat$cell_subtype_formatted) %>%
    as.data.frame.matrix() %>%
    t() %>%
    as.data.frame() %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        colour =  palette_continuous(gray = FALSE)(length(unique(seurat[[]][, "cell_subtype_formatted"]))),
        pct  = FALSE,
        threshold = 1,
        digits = 1
    ) + labs(y = "% Cells")

tab <- table(Assigned = seurat[[]][, "cell_type"], Clusters = Idents(seurat))
pheatmap(log10(tab + 10), color = colorRampPalette(c('white', brewer.pal(9, "Reds")))(10))
```

```{r}
project <- "KO"
saveRDS(seurat, file = file.path(path, paste0(project, ".rds")))
```

```{r}
bcell <- subset(
  seurat, 
  subset = !is.na(seurat$b_cell)
)

saveRDS(bcell, file = file.path(path, paste0(project, "bcell.rds")))
```

## Sample effect
```{r}
table(seurat$Patient)
meta <- seurat@meta.data
meta$cell_id <- rownames(meta)

cells_to_keep <- meta %>%
    group_by(Patient) %>%
    sample_n(size = min(table(seurat$Patient), n()), replace = FALSE) %>%
    pull(cell_id)

seurat_subset <- subset(seurat, cells = cells_to_keep)


DimPlot(
object = seurat_subset,
label = TRUE,
cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
# cols = GimmeMyPlot:::palette_discrete(),
repel = TRUE,
na.value = "white",
reduction = "umap",
split.by = "Patient",
pt.size = 1,
alpha = 0.25
) %>%
theme_sc_dim() + labs(title = NULL, x = NULL, y = NULL)



df_cl <- list.map(
    unique(seurat_subset$Patient),
    f(x) ~ filter(seurat_subset[[]], Patient == x) %>% pull(all_of(snn_cluster)) %>% fct_drop() %>% table()
) %>%
    list.rbind() %>%
    as.data.frame()
    
df_cl %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        colour = col_inds,
        pct  = FALSE,
        digits = 1
    ) + labs(y = "% Cells") +
    geom_hline(yintercept = 50, color = "gray")

df_cl <- list.map(
    unique(seurat$Patient),
    f(x) ~ filter(seurat[[]], Patient == x) %>% pull(all_of(snn_cluster)) %>% fct_drop() %>% table()
) %>%
    list.rbind() %>%
    as.data.frame()
    
df_cl %>%
    t() %>%
    as.data.frame() %>%
    mutate(across(everything(), ~ .x / sum(.x) * 100)) %>%
    plot_bar_2cat(
        count = TRUE,
        stats = FALSE,
        # colour = palette_discrete(),
        colour = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
        legend =  colnames(df_cl),
        pct  = FALSE,
        digits = 1
    ) + labs(y = "% Cells")
```

# Markers

## Predictions from dataset
```{r}
library(readxl)
# fileIn <- file.path(path, "41467_2021_27232_MOESM4_ESM.xlsx")
# name <- "symbol"
fileIn <- file.path(path, "B_cell_subet_Marker_Genes-JCB.xlsx")
name <- "Gene name"

l_bcell_markers <- list.map(
    excel_sheets(fileIn),
    f(i) ~ read_excel(fileIn, sheet = i) %>%
        pull(!!sym(name)) %>%
        na.omit() %>%
        func_format()
) 

p <- list.map(
    l_bcell_markers,
  f(i, j, k) ~ {
      tmp <- list(i) %>%
          AddModuleScore(seurat, features = .)
      stats <- quantile(tmp$Cluster1, probs = c(0.01, 0.99))
      tmp <- subset(tmp, subset = Cluster1 >= stats[1] & Cluster1 <= stats[2])
      VlnPlot(
        object = tmp,
        features = "Cluster1",
        # idents = c(0, 2, 4, 8, 6, 12, 14),
        # cols = palette_discrete(),
        cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(unique(Idents(tmp)))),
        pt.size = 0
      ) + 
          theme_violin_sc() +
        labs(x = NULL, y = NULL, title = str_remove_all(k, " ?vs_all")) + 
        NoLegend()
      }
)

 plot_grid(
    plotlist = p,
    ncol  = 3,
    align = "hv"
)
```

## Cell markers
```{r}
p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ FeaturePlot(
        seurat,
        features = func_format(x) %>% .[. %in% Features(seurat)] %>% head(12),
        min.cutoff = "q10", 
        max.cutoff = "q90",
        cols = c("white", "red"),
        reduction = "umap",
        pt.size = 1,
        alpha = .5,
        combine = FALSE
    )  %>% 
  map(~theme_sc_dim(.) + labs(x = NULL, y = NULL)) %>%
  plot_grid(plotlist = ., align = "hv", ncol = 4, nrow = 3) +
    plot_annotation(
      title = z, 
      theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
    ) 
)

p2 <- list.map(
    l_bcell_markers,
    f(x, y, z) ~ 
        VlnPlot(
            object = seurat,
            features = func_format(x) %>% .[. %in% Features(seurat)] %>% head(12),
            pt.size = 0,
            cols = GimmeMyPlot:::palette_discrete(),
            combine = FALSE
        ) %>%
        map(~. +theme_violin_sc() + labs(y = NULL) + NoLegend()) %>%
        plot_grid(plotlist = ., align = "hv", ncol = 4, nrow = 3) +
        plot_annotation(
            title = z, 
            theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
        ) 
)


p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ DotPlot(
    seurat,
    features = func_format(x) %>% .[. %in% Features(seurat)] %>% head(50) %>% unique(),
    dot.scale = 8,
    scale = FALSE
  ) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(
      x = NULL,
      y = NULL,
      # title = paste("Cluster", x)
      title = z
    )
)
```

```{r}
l_bcell_markers <- list(
  General = c(
    "Pax5", "Ebf1", "Spi1", "Ikzf1", "Ikzf3", # facteurs de transcription clés
    "Pou2f2", "Ets1", "Bach2",                # maintien du programme B
    "Ptprc", "Fnip1", "Hdac5", "Hdac9"
  ),
    
  FrA = c( # B220+CD43+ très précoces
    "Enpep", "Foxp1", "Il7r", "Kit", "Spn"
  ),
  
  FrB = c( # Pro-B précoces
    "Cdkn3", "Dntt", "Igll1", "Irf8", "Lef1", "Rag1", "Rag2", "Vpreb1a", "Vpreb1b", "Il7r"
  ),
  
  FrC = c( # Pro-B tardives (pré-BCR en recombinaison)
     "Polm", "Rag1", "Rag2", "Spib"
  ),
  
  FrD = c( # Pré-B prolifératives
    "Ccna2", "Ccnb1", "Ccnd2", "Ccne1", "Fancg", "Gfi1b",
    "Il2ra", "Mki67", "Pcna", "Polr2a", "Reln", "Top2a"
  ),
  
  FrE = c( # B immatures IgM+
    "Cd19", "Cd79a", "Cd79b", "Ighm", "Igkc", "Iglc2", "Iglc3", "Bcl6"
  ),
  
  FrF = c( # B matures (transit moelle → périphérie)
    "Cd19", "Cd79a", "Cd79b", "Ighm", "Igkc", "Il4i1", "Ms4a1",
    "Nfkbiz", "Notch2", "Ptk2b", "Ptprj", "Tlr9"
  ),
  
  Plasmablast = c(
    "Dnajb9", "Mmp14", "Prdm1", "Sdc1", "Xbp1"
  ),
  
  MemoryB = c(
    "Ahr", "Tnfaip3", "Zfp36l1", "Cd27", "Cd80", "Cd86", "Ms4a1"
  )
  
  # Regulatory<- c(
  #   "Clcf1", "Irf2bp2"
  # )
)
```

## Differentially Expressed Genes
```{r}
cts0 <- AverageExpression(
    seurat,
    group.by = c(snn_cluster),
    # group.by = c(snn_cluster, "Type"),
    assay = assay,
    slot = "data",
    return.seurat = FALSE
) %>% 
  pluck(assay) %>%
  as.data.frame() %>% 
  rename_with(~ str_remove(., "^g")) %>%
  mutate(gene = rownames(.))

cts <- calc_other_cluster_avg(cts0)

# ids <- data.frame(
#     id.1 = c(6, 3, 3, 3, 0, 0, 4, 5, 5),
#     id.2 = c(2, 2, 6, 0, 1, 4, 1, 1, 4)
# )

cluster_marker_genes0 <- list.map(
  levels(Idents(seurat)),
  # seq(nrow(ids)),
  f(x) ~ FindConservedMarkers(
    ident.1 = x,
    # ident.1 = ids[x, 1],
    # ident.2 = ids[x, 2],
    grouping.var = "Patient",
    assay = assay,
    # grouping.var = "Type",
    seurat,
    logfc.threshold = .Machine$double.xmin
  ) %>%
    mutate(
      cluster = x,
      # cluster = paste(ids[x, 2], ids[x, 1], sep = " vs "),
      gene = rownames(.),
      log2FoldChange = rowMeans(select(., ends_with("avg_log2FC")), na.rm = TRUE),
      padj = rowMeans(select(., ends_with("p_val_adj")), na.rm = TRUE),
      `pct.1` = rowMeans(select(., ends_with("pct.1")), na.rm = TRUE),
      `pct.2` = rowMeans(select(., ends_with("pct.2")), na.rm = TRUE)
    ) %>%
  # left_join(select(cts, all_of(c("gene", ids[x, 1], ids[x, 2]))) %>% set_colnames(c("gene", "exp.1", "exp.2")), by = "gene")
  left_join(select(cts, all_of(c("gene", x, paste0("except_", x)))) %>% set_colnames(c("gene", "exp.1", "exp.2")), by = "gene")
)
cluster_marker_genes <- list.rbind(cluster_marker_genes0)

min_p <- cluster_marker_genes %>%
  filter(padj > 0) %>%
  summarise(min_padj = min(padj)) %>%
  pull(min_padj)

cluster_marker_genes$padj[which(cluster_marker_genes$padj == 0)] <- min_p
```

```{r}
fc_threshold  <- log2(1.5)
p_threshold <- .05
pct_threshold <- 0.25

l_dea_raw <- cluster_marker_genes %>%
  mutate(gene_name = gene) %>%
  format_dea(fc_threshold = fc_threshold, p_threshold = p_threshold)
```

### Volcano plots
```{r}
l_dea_raw <- l_dea_raw %>%
    mutate(cluster = factor(cluster, levels = as.character(sort(as.numeric(unique(cluster)))))) %>%
    group_split(cluster, .keep = TRUE) %>%
    set_names(map_chr(., ~as.character(first(.x$cluster)))) %>%
    list.map(
        select(., -cluster) %>%
            mutate(Expression = ifelse(pct.1 > pct_threshold & (pct.1 - pct.2) > pct_threshold, Expression, "ns"))
    )

list.map(
    l_dea_raw, f(x, xx, xxx) ~
        fct_count(x$Expression) %>%
        filter(f!="ns") %>%
        set_colnames(c("Nb DEG", xxx))) %>%
    Reduce(function(x, y) left_join(x, y, by = "Nb DEG"), .) %>%
    bind_rows(summarise(., `Nb DEG` = "Total", across(-`Nb DEG`, sum, na.rm = TRUE))) %>%
    kable0()

l_deg_top <- list.map(
  l_dea_raw,
  f(x) ~ top_genes(
    x,
    fc_threshold = fc_threshold,
    p_threshold = p_threshold,
    n = 15
  )
)

p <- list.map(
    discard(l_dea_raw, function(x) nrow(x) == 0),
    f(x, y, z) ~ {
        volcano_plot(
            x,
            title = z,
            top_genes = l_deg_top[[y]],
            fc_threshold = fc_threshold,
            cex = 1,
            cex_genes = 4,
            p_threshold = p_threshold,
            max.overlaps = 1e3,
            force =1e2
        ) + labs(x = NULL, y = NULL, title = z) +
            theme(legend.position = "none")
    })

plot_grid(plotlist = p, align = "hv", ncol = 4, nrow = 2)

head(l_dea_raw[[1]], n = 10) %>%
    pull("log2FoldChange") %>%
    `^`(2, .) %>%
    plot_bar(label_x = "value", digits = 1, title = "FC")
```

### Dotplots
```{r}
l_dea <- list.map(
    unique(l_dea_raw$cluster),
    f(x) ~ filter(l_dea_raw, cluster == x) %>%
    filter(pct.1 > pct_threshold & (pct.1 - pct.2) > pct_threshold) %>%
  # filter((pct.1 > pct_threshold | pct.2 > pct_threshold) & abs(pct.1 - pct.2) > pct_threshold) %>%
  # TODO: filter cts > 10
      top_genes(return_rank = TRUE, fc_threshold = fc_threshold, p_threshold = p_threshold)
    )
```

```{r}
top10 <- list.map(
  l_dea, 
  f(x) ~ x %>%
    # filter(log2FoldChange > 0) %>%
  slice_head(n = 10)
) %>%
  list.rbind()

last_genes <- top10 %>%
    slice_tail(n = 1)  %>%
    pull(gene)
last_row_numbers <- top10 %>%
    # ungroup() %>%
    filter(!duplicated(gene)) %>%
    mutate(row_num = row_number() %>% rev()) %>%
    # mutate(row_num = row_number()) %>% # if pseudo
    group_by(cluster) %>%
    slice_tail(n = 1) %>%
    pull(row_num) %>%
    sort()

DoHeatmap(
    seurat,
    features = top10$gene,
    # group.by = "cell_type_na",
    group.colors = palette_discrete(),
    slot = "data"
) + 
    NoLegend() + 
    scale_fill_gradientn(
        colors = c("blue", "white", "red"),
        na.value = "black"
    ) +
    geom_hline(yintercept = last_row_numbers[-1] - 0.5, color = "black", linetype = "solid", linewidth = 1)


last_row_numbers <- top10 %>%
    # ungroup() %>%
    filter(!duplicated(gene)) %>%
    mutate(row_num = row_number()) %>%
    group_by(cluster) %>%
    slice_tail(n = 1) %>%
    pull(row_num) %>%
    sort()

DotPlot(
    seurat, 
    features = unique(pull(top10, "gene")), 
    dot.scale = 8,
    scale = FALSE
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL) +
    geom_vline(xintercept = last_row_numbers[-1] + 0.5)

cluster_marker_genes %>% 
    filter(cluster == 0) %>%
  head(15) %>%
  mutate(padj = format(padj, digits = 1, scientific = TRUE)) %>%
  mutate(FC = 2^abs(log2FoldChange) * sign(log2FoldChange)) %>%
   set_rownames(pull(., gene)) %>%
  select(FC, padj) %>%
  kable0()
```

```{r}
top50 <- list.map(
  l_dea, 
  f(x) ~ x %>%
    # filter(log2FoldChange > 0) %>%
  slice_head(n = 50)
) %>%
  list.rbind()

p2 <- list.map(
  unique(pull(top50, cluster)),
  f(x) ~ DotPlot(
    seurat,
    features = filter(top50, cluster == x) %>% pull("gene") %>% unique(),
    dot.scale = 8,
    scale = FALSE
  ) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(
      x = NULL,
      y = NULL,
      # title = paste("Cluster", x)
      title = x
    )
)
```

```{r marker_cell}
RidgePlot(
  seurat_integrated, 
  features = top_genes2[[1]], 
  ncol = 4,
  cols = col_clus
)
VlnPlot(
  seurat_integrated, 
  features = top_genes2[[1]],
  pt.size = 0,
  cols = col_clus
)
VlnPlot(
  seurat_integrated, 
  features = top_genes2[[1]],
  pt.size = 0, 
  split.by = "Patient", 
  split.plot = TRUE
)
# VlnPlot(seurat_integrated, features = top_genes[[1]], slot = 'counts', log = TRUE)
FeaturePlot(
  seurat_integrated,
  features = top_genes2[[1]],
  min.cutoff = "q10", 
  max.cutoff = "q90",
  label = TRUE,
  cols = c("grey", "red")
)

DoHeatmap(
  subset(seurat_integrated, downsample = 100),
  features = top_genes2[[2]],
  group.colors = col_clus,
  group.by = "cell_type"
  ) + scale_fill_gradientn(
    colors = c("blue", "white", "red"), 
    na.value = "black"
  )
DotPlot(
  seurat_integrated, 
  features = top_genes2[[2]], 
  cols = c('blue', 'red'), 
  dot.scale = 8,
  split.by = "Patient"
  ) +
  RotatedAxis()+
  theme_custom()
```

## Condition markers
```{r feature.heatmaps, eval = FALSE}
FeaturePlot(
  seurat_integrated, 
  features = head(dgea, n = 3) %>% rownames(), 
  split.by = "Patient", 
  min.cutoff = "q10",
  max.cutoff = "q90",
  # cols = c("grey", "red"),
  reduction = 'umap'
)

VlnPlot(
    seurat_integrated, 
    features = head(top_genes, 4) %>% rownames(),
    pt.size = 0, 
    group.by = "cell_type",
    split.by = "Patient", 
    split.plot = TRUE,
    ncol = 2
)

library(ComplexHeatmap)
rowAnnotation(
  "HV" = anno_barplot(top_genes$pct.1),
  "AOSD"= anno_barplot(top_genes$pct.2),
  width = unit(10, "cm")
) %>%
  Heatmap(
    top_genes$avg_log2FC,
    name = "Log2FC",
    cluster_rows = TRUE,
    row_labels = rownames(top_genes),
    right_annotation = .,
    width = unit(1, "cm")
    )

theme_set(theme_cowplot())
pop_seurat <- subset(seurat_integrated, idents = "G1")
Idents(pop_seurat) <- "Patient"
avg_pop <- as.data.frame(log1p(AverageExpression(pop_seurat, verbose = FALSE)$RNA))
avg_pop$gene <- rownames(avg_pop)
p <- ggplot(avg_pop, aes(S01004A, S01013A)) + 
  geom_point() + 
  ggtitle("G1")
LabelPoints(p, points = top_genes %>% head(15) %>% rownames(), repel = TRUE, color = "red")
```

# Pseudo-bulk
```{r scatterplots}
# seurat_integrated@meta.data[, "sample"] <- paste0(
#   seurat_integrated@meta.data[, "Patient"],
#   seurat_integrated@meta.data[, "Type"],
# )
# seurat$cell_type <- seurat[[]][, clusters[nb_optimal]]
# seurat$cell_type <- Idents(seurat) %>% 
  # fct_relevel(as.character(0:9))
cts <- AggregateExpression(
  seurat,
  group.by = c("predicted.labels", "Patient"),
  assay = "RNA",
  slot = "counts",
  return.seurat = FALSE
)

# bulk.mono.de <- FindMarkers(
#   object = cts,
#   ident.1 = "G1_S01004A",
#   ident.2 = "G1_S01013A",
#   test.use = "DESeq2"
# )
# head(bulk.mono.de, n = 15)

cts <- as.data.frame(cts$RNA)
cts <- list.map(
  levels(seurat@meta.data[, "predicted.labels"]),
  f(x) ~{
    select(cts, starts_with(paste0("g", x, "_"))) %>% 
      set_colnames(colnames(.) %>% str_remove_all(paste0("g", x, "_")))
  }
) %>% set_names(levels(seurat@meta.data[, "predicted.labels"]))

saveRDS(cts, file.path(path, "l_cts.rds"))

pct <- pct_by_ident_type(seurat)
saveRDS(pct, file.path(path, "sc_metadata.rds"))
```

```{r}
l_deg0 <- readRDS(file.path(path, "l_deg.rds"))
l_deg <- discard(l_deg0, ~ nrow(.x) == 0)
top10 <- list.map(l_deg,  f(x, xx, xxx) ~head(x, 10) %>% mutate(cluster = xxx)) %>%
    list.rbind() %>%
    rename(gene = "gene_name")
l_bcell_markers <- list.map(l_deg, unique(.) %>% head(12) %>% pull(Gene))
DotPlot(
    seurat, 
    features = unique(pull(top10, "gene")), 
    dot.scale = 8,
    scale = FALSE,
    group.by = "cell_type",
    split.by = "Type",
    cols = GimmeMyPlot:::palette_discrete()
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL) +
    geom_vline(xintercept = last_row_numbers[-last_id] + 0.5) +
    geom_hline(yintercept = seq(2, (length(l_deg0) -1) * 2, by = 2) + 0.5)


l_bcell_markers <- list.map(l_deg, unique(.) %>% pull(Gene))
p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ FeaturePlot(
        seurat,
        features = func_format(x) %>% .[. %in% Features(seurat)] %>% head(4),
        min.cutoff = "q10", 
        max.cutoff = "q90",
        cols = c("white", "red"),
        split.by = "Type",
        reduction = "umap",
        pt.size = 1,
        alpha = .5,
        combine = FALSE
    )  %>% 
  map(~theme_sc_dim(.) + labs(x = NULL, y = NULL)) %>%
  plot_grid(plotlist = ., ncol = 4) +
    plot_annotation(
      title = z, 
      theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
    ) 
)

p2 <- list.map(
  l_bcell_markers,
  f(x, y, z) ~ 
VlnPlot(
    object = seurat,
    features = func_format(x) %>% .[. %in% Features(seurat)] %>% head(12),
    pt.size = 0,
    group.by = "cell_type",
    split.by = "Type",
    cols = cols,
    combine = FALSE
) %>%
map(~. +theme_violin_sc() + labs(y = NULL) + NoLegend()) %>%
  plot_grid(plotlist = ., ncol = 4, nrow = 3) +
    plot_annotation(
      title = z, 
      theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
    ) 
)
# i = 2; plot_grid(plotlist = p [(12*(i-1)+1):(12*i)], ncol = 4, nrow = 3)


list.map(
    split(cell_markers_bcell, ceiling(seq_along(cell_markers_bcell)/4)),
f(x) ~ 
FeaturePlot(
    seurat,
    features = x,
    min.cutoff = "q10", 
    max.cutoff = "q90",
    cols = c("white", "red"),
    split.by = "Type",
    reduction = "umap",
    pt.size = 1,
    alpha = .5,
    combine = FALSE
)  %>% 
    map(~theme_sc_dim(.) + labs(x = NULL, y = NULL))  %>%
    plot_grid(plotlist = ., ncol = 4)
)

top10 <- l_bcell_markers %>%
  list.map(f(x, y, z) ~  x[x %in% Features(seurat)] %>% head(10) %>% data.frame(gene = ., cluster = y)) %>% 
  list.rbind() %>%
  filter(!duplicated(gene))

DotPlot(
    seurat, 
    features = list.map(l_bcell_markers, .[. %in% Features(seurat)] %>% head(10)) %>% unlist() %>% unique(), 
    dot.scale = 8,
    scale = FALSE,
    group.by = "cell_type",
    split.by = "Type",
    cols = GimmeMyPlot:::palette_discrete()
) +
    RotatedAxis() +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL, y = NULL) +
    geom_vline(xintercept = last_row_numbers[-last_id] + 0.5) +
    # geom_vline(xintercept = seq(10, 60, 10) + 0.5) +
    geom_hline(yintercept = seq(2, 12, by = 2) + 0.5)

list.map(
    top50,
    f(i, ii, z) ~
    list.map(
        func_format(i) %>% .[. %in% Features(seurat)] %>% head(4*3) %>% split(., ceiling(seq_along(.)/4)),
    f(x) ~ 
        FeaturePlot(
            seurat,
            features = x,
            min.cutoff = "q10", 
            max.cutoff = "q90",
            cols = c("white", "red"),
            split.by = "Type",
            reduction = "umap",
            pt.size = 1,
            alpha = .5,
            combine = FALSE
        )  %>% 
        map(~theme_sc_dim(.) + labs(x = NULL, y = NULL))  %>%
        plot_grid(plotlist = ., ncol = 4) +
        plot_annotation(
            title = z, 
            theme = theme(plot.title = element_text(face = "bold", size = 18, hjust = 0.5))
        )
)
) -> p
```

```{r }
cts0 <- cts[[1]]
colData <- data.frame(samples = colnames(cts0)) %>%
  mutate(condition = str_extract(samples, unique(seurat$Type) %>% paste(collapse = "|")) %>% as.factor()) %>%
  column_to_rownames(var = 'samples')

dds <- DESeqDataSetFromMatrix(
  countData = cts0,
  colData = colData,
  design = ~ condition
) %>%
   .[rowSums(counts(.) >= 10) >= 3, ] %>%
  DESeq(quiet = TRUE)


resultsNames(dds)[2]
res <- results(dds, name = resultsNames(dds)[2])

# genes.to.label <- NULL
# p <- lapply(
#   c("S01004A", "S01013A"),
#   function(x) {
#   CellScatter(aggregate_seurat, paste0("G0_", x), paste0("G1_", x), highlight = genes.to.label) %>%
#   LabelPoints(points = genes.to.label, repel = TRUE)
# })
```

```{r}
colours <- get_colors()[c(3, 1, 4:5)]
list.map(tmpp0, f(x, y) ~data.frame(Type = colours[y], val = x)) %>% list.rbind() %>% set_rownames(rownames(.) %>% str_remove_all("A\\d{2}PB\\d")) -> cou
plot_bar(select(cou, 2), colour = pull(cou, 1), label_x = "none", ratio = 10) + theme(axis.text.y = element_text(colour = "gray"), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), panel.grid.major.y = element_line(colour = "grey", linewidth = 0.5), panel.grid.major.x = element_blank()) + ggtitle(NULL)
```

# Trajectory analysis

## Monocle
```{r}
library(SeuratWrappers)
cds <- as.cell_data_set(seurat)
```

```{r}
library(monocle3)
start_cluster <- 6
names(cds@clusters$UMAP$partitions) <- colnames(cds)
root_cells <- colnames(cds)[cds@clusters$UMAP$clusters == start_cluster]

cds <- learn_graph(cds, use_partition = FALSE) %>%
  order_cells(root_cells = root_cells)

cds@principal_graph_aux@listData$UMAP$pseudotime <- max(pseudotime(cds), na.rm = TRUE) - pseudotime(cds)
```

```{r}
plot_cells(
    cds,
    color_cells_by = "pseudotime",
    label_root = FALSE,
    label_leaves = FALSE,
    label_branch_points = FALSE,
    trajectory_graph_color = "black",
    trajectory_graph_segment_size = 1.5,
    cell_size = 1,
    alpha = .15
) %>% theme_sc_dim() + 
    scale_color_gradientn(
      colors = pal_sc, 
      name = "Pseudotime"
    )

# seurat <- as.Seurat(cds, assay = "integrated")
```

## Slingshot
```{r}
library("slingshot")

sce <- as.SingleCellExperiment(seurat)
reducedDims(sce)$UMAP <- Embeddings(seurat, reduction = "umap") -> reduc_coords

sce <- slingshot(
  sce,
  start.clus = start_cluster,
  clusterLabels = "cell_type",
  reducedDim = "UMAP"
)

curves <- slingCurves(sce)
lineage_probs <- list.map(curves, .$w) %>%
  list.cbind() %>%
  max.col(ties.method = "first")

pseudotime <- slingPseudotime(sce) %>%
  .[cbind(seq_len(nrow(.)), lineage_probs)]

plot(
  reduc_coords,
  # col = GimmeMyPlot:::palette_discrete()[sce$cell_type],
  col = colorRampPalette(pal_sc)(100)[cut(pseudotime, breaks = 100)],
  pch = 16,
  asp = 1
)
# lines(SlingshotDataSet(sce), lwd = 4, col = "black")
lines(SlingshotDataSet(sce), lwd = 4, type = 'lineages', col = "black",  show.constraints = TRUE)

p <- as.data.frame(reduc_coords) %>%
  mutate(pseudotime = pseudotime) %>%
  ggplot(aes(x = umap_1, y = umap_2, colour = pseudotime)) +
  geom_point(alpha = 0.8)

p <- reduce(
  names(curves),
  .f = function(plot, lineage_name) {
    plot + geom_path(
      data = as.data.frame(curves[[lineage_name]]$s),
      aes(x = umap_1, y = umap_2),
      size = 1.2,
      color = "black",
      inherit.aes = FALSE
    )
  },
  .init = p
)
theme_sc_dim(p)  + 
    scale_color_gradientn(
      colors = pal_sc, 
      name = "Pseudotime"
    )
```

# Transfer learning
```{r}
wt@meta.data[, "labels"] <- Idents(wt)
# wt@assays$SCT@scale.data <- matrix()
# ko@assays$SCT@scale.data <- matrix()
seurat <- merge(wt, y = ko)
```

```{r}
anchors <- FindTransferAnchors(
    reference = wt,
    query = seurat,
    # normalization.method = "SCT",
    reference.reduction = "pca",
    # reference.reduction = integration_method,
    dims = seq(n_dim)
)
```

```{r, eval = FALSE}
cl_trans <- TransferData(
    anchorset = anchors,
    dims = seq(n_dim),
    # weight.reduction = KO[[integration_method]],
    refdata = Idents(wt)
  )
seurat <- AddMetaData(seurat, metadata = cl_trans)
```

```{r, eval = FALSE}
wt2 <- RunUMAP(
  wt,
  dims = seq(n_dim),
  # reduction = integration_method,
  reduction = "pca",
  return.model = TRUE
)
```

```{r}
seurat <- MapQuery(
    anchorset = anchors,
    query = seurat,
    reference = wt,
    refdata = list(labels = "labels"),
    # reference.reduction = integration_method,
    reduction.model = "umap"
)
```

```{r}
plot_sc_violin(
    seurat,
    features = paste0("prediction.score.", levels(Idents(wt))),
    cols = GimmeMyPlot:::palette_discrete(),
    nrow = 3
)
```

```{r}
type <- NULL
seurat$predicted.labels<- factor(seurat$predicted.labels, levels = 0: (length(unique(seurat$predicted.labels)) - 1))
Idents(seurat) <- "predicted.labels"
DimPlot(
    object = seurat, 
    label = TRUE,
    # cols = GimmeMyPlot:::palette_continuous(gray = FALSE)(length(levels(Idents(seurat)))),
    cols = GimmeMyPlot:::palette_discrete(),
    repel = TRUE,
    na.value = "white",
    split.by = "Type",
    # group.by = ,
    reduction = "ref.umap",
    pt.size = 1,
    alpha = 0.25
) %>% 
    theme_sc_dim() + ggtitle(type) + NoLegend()
```

## Session information
```{r end, echo = FALSE}
sessionInfo()
```
